---
title: "Describing the NVI and Comparison the the NVI to the NDI/SVI: Results and Figures"
author: 
- "Stephen P. Uong" 
- "Contributors: Jiayi Zhou, Jeanette A. Stingone"
date: "10/28/2021"
output: 
  html_document: 
    toc: true
    toc_float: true
---

We developed a Neighborhood Vulnerability Index (NVI) to measure vulnerability to environmental pollution. We applied the Toxicological Prioritization Index (ToxPi) to integrate data across different domains, characterize neighborhood vulnerability to environmental pollution in New York City (NYC), and determine if sources of vulnerability varied across neighborhoods.

In this document, we have included code we used to generate results and figures that we used when evaluating the NVI and comparing the NVI to the Neighborhood Deprivation Index (NDI)/ Social Vulnerability Index (SVI).

### 1. Setup
We set the working directory to one folder up from the RMarkdown file.
```{r, setup}
knitr::opts_knit$set(root.dir = '..') 
```

### 2. Loading Libraries
```{r, libraries}
library(tidyverse)
library(rio)
library(skimr)
library(janitor)
library(cluster)
library(factoextra)
# library(Amelia)
# library(janitor)
# library(cluster)
# Mapping
library(sf)
library(ggpubr)
library(ggsn)
library(ggspatial)
library(nycgeo)
# Plots
library(ggpubr)
library(colorBlindness)
library(ggpattern) # Ref: https://coolbutuseless.github.io/package/ggpattern/index.html

```


### 3: Importing the Index Generated from the ToxPi GUI, NVI features, NDI, and SVI
```{r}
setwd('C:/Users/steph/OneDrive - cumc.columbia.edu/Phi/My Courses/Columbia/Work/Research - Jeanette/projects/covid-vulnerability-index/comms/git nvi')

### IMPORT: Cleaned NVI data with the NVI, NVI features, and NVI clusters
tract_final <- readRDS(file = "data/processed/nvi_final.rds")
### IMPORT: NYC specific Census tract boundaries
tract_nyc_spatial <- nycgeo::nyc_boundaries(geography = "tract") %>% 
  dplyr::transmute(Tract_FIPS = geoid)
tract_nyc_county <- nycgeo::nyc_boundaries(geography = "tract") %>% 
  dplyr::as_tibble() %>% 
  dplyr::select(-geometry) %>% 
  dplyr::transmute(Tract_FIPS = geoid,
                   county = county_name)

### IMPORT: Exclusions List
tract_exclusions_list <- import("data/processed/EXCLUSION_LIST_20210628.csv")
tract_exclusions_list_id <- tract_exclusions_list %>% 
  dplyr::filter(flag_exclude_FINAL == 0) %>% 
  dplyr::transmute(SID = as.character(GEOID))

### IMPORT: NDI, SVI
ndi <- import("data/processed/NDI.NYC.2086t.csv")
svi <- import("data/raw/svi.csv")

ndi_clean <- ndi %>% dplyr::mutate(Tract_FIPS = GEOID %>% as.character() %>% str_pad(11,side="left",pad="0")) %>% dplyr::rename(ndi = NDI_score) %>% dplyr::select(Tract_FIPS, ndi) %>% 
  dplyr::left_join(tract_nyc_county, by = "Tract_FIPS")
svi_clean <- svi %>% dplyr::mutate(Tract_FIPS = FIPS %>% as.character() %>% str_pad(11,side="left",pad="0")) %>% dplyr::rename(svi = RPL_THEMES) %>% dplyr::select(Tract_FIPS, svi) %>% 
  dplyr::filter(svi != -999) %>% 
  dplyr::left_join(tract_nyc_county, by = "Tract_FIPS") %>% 
  dplyr::inner_join(tract_exclusions_list_id %>% dplyr::transmute(Tract_FIPS = SID), by = "Tract_FIPS")

## IMPORT: Race/ethnicity sensitivity analysis
tract_toxpi_pre_raceeth_no_olap <- import('data/processed/toxpi/nvi_toxpi_raceeth_no_olap_results.csv')
tract_toxpi_pre_raceeth_olap <- import('data/processed/toxpi/nvi_toxpi_raceeth_olap_results.csv')
tract_toxpi_pre_raceeth_hsub <- import('data/processed/toxpi/nvi_toxpi_raceeth_hsub_results.csv')

```

### Create spatial objects
```{r}
# Spatial: crs = "+init=epsg:2263"
  # Included
tract_final_spatial <- tract_nyc_spatial %>% 
  dplyr::inner_join(tract_final %>% dplyr::mutate(Tract_FIPS = as.character(Tract_FIPS) %>% trimws()), by = "Tract_FIPS") # Checked: did not exclude any tracts, 8/11
  # Excluded
tract_exclude_spatial <- tract_nyc_spatial %>% 
  dplyr::anti_join(tract_final %>% dplyr::mutate(Tract_FIPS = as.character(Tract_FIPS) %>% trimws()), by = "Tract_FIPS") 
```


# FIGURES
## FIGURE 1: Domains and subdomains
```{r mapping, include=FALSE}
#### CREATE: Functions to create maps
gen_fig1_main <- function(dfin, dfexcl, fill_var, fill_color, fill_label){
  plot_choro <- ggplot() + 
    geom_sf(data = dfin, aes(fill = !!sym(fill_var)), color = NA) + 
    geom_sf(data = dfexcl, color = "gray30", fill = "gray30") + 
    theme_minimal() + # Alt: ggthemes::theme_map()
    ggsn::scalebar(data = dfin, dist = 20, dist_unit = "mi", height = 0.01, st.size = 5, st.dist = 0.05, transform = FALSE, model = 'WGS84', location = "bottomright") + # scale bar
    ggsn::north(data = dfin, location = "topleft", symbol = 10, scale = 0.10) + # north arrow, symbol = 16, scale = 0.17
        # Alt: ggspatial::annotation_north_arrow(pad_x = unit(2.5, "cm"), pad_y = unit(-1.75, "cm")) +
        # Alt: ggspatial::annotation_scale(plot_unit = "mi", pad_x = unit(7.5, "cm"), pad_y = unit(-1, "cm")) #bgcolor("lightcyan") + #snow2 
    labs(fill = fill_label) + 
    scale_fill_gradient(low = "whitesmoke", high = fill_color) + # na.value = "gray30"
    theme(legend.title = element_text(face = "bold", size = 15), # plot.title = element_text(hjust = 0.75, face = "bold", size = 15),
          legend.text = element_text(size = 12),
          legend.spacing.x = unit(0.5, 'cm'),
          legend.position = "right",
          axis.line = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank(),
          panel.grid = element_blank()) # plot.caption = element_text(hjust = 0, size = 10, face = "italic"), legend.position = c(1,0.5), plot.background = element_rect(fill = "lightcyan")
  return(plot_choro)
}
gen_fig1_part <- function(dfin, dfexcl, fill_var, fill_color, fill_label, fill_label_just, fill_increment){
  plot_choro <- ggplot() + 
    geom_sf(data = dfin, aes(fill = !!sym(fill_var)), color = NA) + 
    geom_sf(data = dfexcl, color = "gray30", fill = "gray30") + 
    theme_minimal() + 
    scale_fill_gradient(low = "whitesmoke", high = fill_color, breaks = seq(0, 1, fill_increment)) + 
    labs(fill = fill_label) + 
    theme(legend.title = element_text(face = "bold", size = 14, vjust = fill_label_just),  # vertically center legend with its label - does not work well
          legend.text = element_text(size = 10),
          legend.spacing.x = unit(0.5, 'cm'),
          legend.position = "right",
          axis.line = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank(),
          panel.grid = element_blank())
  return(plot_choro)
}

### CREATE: Maps
  # Overall Index
fig1_toxpi <- gen_fig1_main(dfin = tract_final_spatial, dfexcl = tract_exclude_spatial, fill_var = "nvi", fill_color = "#D55E00", fill_label = "Neighborhood\nVulnerability Index")
  # Domain Scores
fig1_demo <- gen_fig1_part(dfin = tract_final_spatial, dfexcl = tract_exclude_spatial, fill_var = "score_demo", fill_color = "#0072B2", fill_label = "Demographics Score", fill_label_just = 0.5, fill_increment = 0.1)
fig1_econ <- gen_fig1_part(dfin = tract_final_spatial, dfexcl = tract_exclude_spatial, fill_var = "score_economic", fill_color = "#009E73", fill_label = "Economic\nIndicators Score", fill_label_just = 0.5, fill_increment = 0.2)
fig1_resd <- gen_fig1_part(dfin = tract_final_spatial, dfexcl = tract_exclude_spatial, fill_var = "score_residential", fill_color = "#E69F00", fill_label = "Residential Density\nand Characteristics\nScore", fill_label_just = 4, fill_increment = 0.2)
fig1_hlth <- gen_fig1_part(dfin = tract_final_spatial, dfexcl = tract_exclude_spatial, fill_var = "score_healthstatus", fill_color = "#CC79A7", fill_label = "Health Behavior,\nOutcomes,\nPrevention Practices,\nand Access Score", fill_label_just = 4, fill_increment = 0.2)
## MERGE: Maps
  # Alt: gridExtra::grid.arrange(fig1_toxpi, fig1_demo, fig1_econ, fig1_resd, fig1_hlth, ncol = 4, nrow = 2, layout_matrix = rbind(c(1,1,2,3), c(1,1,4,5))) 
  # Ref: https://www.datanovia.com/en/lessons/combine-multiple-ggplots-into-a-figure/
fig1_part1 <- ggarrange(fig1_toxpi, fig1_demo, nrow = 2, labels = c("A","B"), align = "v", font.label = list(size = 25)) + theme(plot.margin = margin(t = 0, r = 0, b = 1, l = 0, "cm"))
fig1_part2 <- ggarrange(fig1_econ, fig1_resd, nrow = 2, labels = c("C","D"), align = "v", font.label = list(size = 25)) + theme(plot.margin = margin(t = 0, r = 0, b = 1, l = 0, "cm"))
fig1_part3 <- ggarrange(fig1_hlth, labels = "E", font.label = list(size = 25)) + theme(plot.margin = margin(t = 0, r = 0, b = 1, l = 0, "cm"))

fig1_part2 <- ggarrange(fig1_econ, ggarrange(fig1_resd, labels = "D", font.label = list(size = 25)), nrow = 2, labels = "C", align = "v", font.label = list(size = 25)) +
  theme(plot.margin = margin(t = 0, r = 0, b = 1, l = 0, "cm"))


## EXPORT
fig1_part1
ggsave('figs/fig1_part1.png', width = 10, height = 15) # pixels: 1200 x 900, width = 15, height = 12
fig1_part2
ggsave('figs/fig1_part2.png', width = 10, height = 15) 
fig1_part3
ggsave('figs/fig1_part3.png', width = 10, height = 8) 


```

## FIGURE 2. MAP: NVI, NDI, SVI
```{r}
#### CREATE: Functions to create maps
gen_fig2_main <- function(dfin, dfexcl, fill_var, fill_color, fill_label){
  plot_choro <- ggplot() + 
    geom_sf(data = dfin, aes(fill = !!sym(fill_var)), color = NA) + 
    geom_sf(data = dfexcl, color = "gray30", fill = "gray30") + 
    theme_minimal() + # Alt: ggthemes::theme_map()
    ggsn::scalebar(data = dfin, dist = 20, dist_unit = "mi", height = 0.01, st.size = 5, st.dist = 0.05, transform = FALSE, model = 'WGS84', location = "topleft") + # scale bar, bottomright
    ggsn::north(data = dfin, location = "bottomright", symbol = 10, scale = 0.10) + # north arrow, symbol = 16, scale = 0.17, topleft
        # Alt: ggspatial::annotation_north_arrow(pad_x = unit(2.5, "cm"), pad_y = unit(-1.75, "cm")) +
        # Alt: ggspatial::annotation_scale(plot_unit = "mi", pad_x = unit(7.5, "cm"), pad_y = unit(-1, "cm")) #bgcolor("lightcyan") + #snow2 
    labs(fill = fill_label) + 
    #scale_fill_gradient(low = "whitesmoke", high = fill_color) + # na.value = "gray30"
    scale_fill_manual(values = fill_color, guide = guide_legend(direction = "horizontal", title.position = "top")) + # cannot seem to get title to center: title.hjust = 0.5
    theme(legend.title = element_text(face = "bold", size = 15), # plot.title = element_text(hjust = 0.75, face = "bold", size = 15),
          legend.text = element_text(size = 12),
          legend.spacing.x = unit(0.5, 'cm'),
          legend.position = "bottom",
          legend.box = "vertical",
          axis.line = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank(),
          panel.grid = element_blank())  # plot.caption = element_text(hjust = 0, size = 10, face = "italic"), legend.position = c(1,0.5), plot.background = element_rect(fill = "lightcyan")
    #guides(size = guide_legend(title.position = "top", title.hjust = 0.5))
  return(plot_choro)
}
gen_fig2_part <- function(dfin, dfexcl, fill_var, fill_color, fill_label, fill_label_just, fill_increment){
  plot_choro <- ggplot() + 
    geom_sf(data = dfin, aes(fill = !!sym(fill_var)), color = NA) + 
    geom_sf(data = dfexcl, color = "gray30", fill = "gray30") + 
    theme_minimal() + 
    scale_fill_manual(values = fill_color, guide = guide_legend(direction = "horizontal", title.position = "top")) +
    labs(fill = fill_label) + 
    theme(legend.title = element_text(face = "bold", size = 14, vjust = fill_label_just),  # vertically center legend with its label - does not work well
          legend.text = element_text(size = 10),
          legend.spacing.x = unit(0.5, 'cm'),
          legend.position = "bottom",
          axis.line = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank(),
          panel.grid = element_blank()) 
  return(plot_choro)
}

### PREP DATA
fig2_sf <- tract_final_spatial %>% 
  dplyr::select(Tract_FIPS, nvi) %>% 
  dplyr::left_join(ndi_clean, by = "Tract_FIPS") %>% 
  dplyr::left_join(svi_clean, by = "Tract_FIPS") %>% 
  dplyr::mutate(nvi_q = ntile(nvi, 4) %>% as.factor(),
                ndi_q = ntile(ndi, 4) %>% as.factor(),
                svi_q = ntile(svi, 4) %>% as.factor()) 

### CREATE: Maps for NVI, NDI, SVI
red_fill <- c('#F6DECC', '#E59E66', '#d55e00', '#954100') # "#fee5d9","#fcae91","#fb6a4a","#cb181d"
fig2_toxpi <- gen_fig2_main(dfin = fig2_sf, dfexcl = tract_exclude_spatial, fill_var = "nvi_q", fill_color = red_fill, fill_label = "Neighborhood Vulnerability Index Quartile") 
fig2_ndi <- gen_fig2_part(dfin = fig2_sf, dfexcl = tract_exclude_spatial, fill_var = "ndi_q", fill_color =  red_fill, fill_label = "Neighborhood Deprivation Index Quartile", fill_label_just = 0.5, fill_increment = 0.1) 
fig2_svi <- gen_fig2_part(dfin = fig2_sf, dfexcl = tract_exclude_spatial, fill_var = "svi_q", fill_color =  red_fill, fill_label = "Social Vulnerability Index Quartile", fill_label_just = 0.5, fill_increment = 0.2) 
## MERGE: Maps
fig2 <- ggarrange(fig2_toxpi, fig2_ndi, fig2_svi, ncol = 3, labels = c("A","B","C")) + theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0, "cm"))
# fig2 <- ggarrange(fig2_toxpi, ggarrange(fig2_ndi, fig2_svi, nrow = 2, labels = "B", font.label = list(size = 25)), ncol = 2, labels = "A", font.label = list(size = 25)) + theme(plot.margin = margin(t = 0, r = 0, b = 2, l = 0, "cm"))
fig2
ggsave('figs/fig2.png', width = 13, height = 6) # pixels: 1200 x 900 width = 13, height = 12

```

## FIGURE 3 / SUPPLEMENTAL FIGURE 2. Scatter plots
```{r}
### CREATE: Function to generate scatterplots
  # Ref, stat_cor: https://www.rdocumentation.org/packages/ggpubr/versions/0.4.0/topics/stat_cor
gen_fig3_scatterplot <- function(datain, var, var_label, var_breaks){
  plot_scatter <- ggplot(datain, aes(x = nvi, y = !!sym(var), color = nvi_domain)) +
    geom_point(alpha = 0.5, size = 0.8) + #  geom_smooth(method = lm) +
    ggpubr::stat_cor(aes(label = ..r.label..), method = "spearman", label.x.npc = 0.6, label.y.npc = 0.25, p.accuracy = 0.001, size = 6, color = "black") + # "~`,`~" # label.x.npc = 0.7, size = 4
    # ggpubr::stat_cor(aes(label = ..p.label..), method = "spearman", label.x.npc = 0.7, label.y.npc = 0.1, p.accuracy = 0.001, size = 4, color = "black") + # "~`,`~"
    facet_grid(county ~ nvi_domain) + 
    scale_color_manual(values = c("#D55E00","#0072B2","#009E73","#E69F00","#CC79A7")) +
    scale_x_continuous(breaks = seq(-1, 0.8, 0.2), minor_breaks = seq(-1, 0.9, 0.1)) +
    scale_y_continuous(breaks = seq(-2, 6, var_breaks), minor_breaks = seq(-2, 6, var_breaks/2)) +
    theme_bw() +
    labs(x = "Neighborhood Vulnerability Index", y = var_label) +
    theme(legend.position = "none",
          axis.title.x = element_text(size = 20, margin = margin(t = 10)),
          axis.title.y = element_text(size = 20, margin = margin(r = 10)),
          axis.text.x = element_text(size = 13),
          axis.text.y = element_text(size = 13),
          strip.text.x = element_text(size = 15),
          strip.text.y = element_text(size = 15))
  return(plot_scatter)
}

### DATA PREP
fig3_nvi_df <- tract_final %>% 
  dplyr::select(Tract_FIPS, county, nvi, score_demo, score_economic, score_residential, score_healthstatus) %>% 
  tidyr::gather(key = nvi_domain, value = nvi, nvi, score_demo, score_economic, score_residential, score_healthstatus) %>% 
  dplyr::mutate(nvi_domain = case_when(nvi_domain == "nvi" ~ "Overall:\nNeighborhood\nVulnerability Index",
                                       nvi_domain == "score_demo" ~ "Domain:\nDemographic Score",
                                       nvi_domain == "score_economic" ~ "Domain:\nEconomic Score",
                                       nvi_domain == "score_residential" ~ "Domain:\nResidential Score",
                                       nvi_domain == "score_healthstatus" ~ "Domain:\nHealth Status Score") %>% 
                                       factor(levels = c("Overall:\nNeighborhood\nVulnerability Index","Domain:\nDemographic Score","Domain:\nEconomic Score","Domain:\nResidential Score","Domain:\nHealth Status Score")))
### MERGE with NDI/SVI data
fig3_df <- fig3_nvi_df %>%
  dplyr::mutate(Tract_FIPS = as.character(Tract_FIPS)) %>% 
  dplyr::left_join(ndi_clean %>% dplyr::select(-county), by = "Tract_FIPS") %>% 
  dplyr::left_join(svi_clean %>% dplyr::select(-county), by = "Tract_FIPS")
### CREATE: Plots
fig3_ndi <- gen_fig3_scatterplot(datain = fig3_df, var = "ndi", var_label = "Neighborhood Deprivation Index", var_breaks = 2) # NDI
figs2_svi <- gen_fig3_scatterplot(datain = fig3_df, var = "svi", var_label = "Social Vulnerability Index", var_breaks = 0.2) # SVI
### EXPORT: Plots
fig3_ndi
ggsave('figs/fig3.png', width = 15, height = 12)
figs2_svi
ggsave('figs/figs2.png', width = 15, height = 12)


```

## FIGURE 4 - PART B (CLUSTER MAP + CLUSTER PROFILE)
```{r}
## CHECK WHICH CLUSTERS TRACTS WERE IN
tract_final %>% filter(Tract_FIPS %in% c('36047051900','36081067100')) %>% select(Tract_FIPS, county, nvi, nvi_cluster, score_demo, score_economic, score_residential, score_healthstatus) # Queens tract: 4, Kings tract: 6
#View where on map: https://popfactfinder.planning.nyc.gov/#11.95/40.72281/-73.95688
# library(leaflet)
# tract_final_spatial_tempproj <- st_transform(tract_final_spatial, crs = '+proj=longlat +datum=WGS84')
# tract_final_spatial_tempproj_subset <- tract_final_spatial_tempproj %>% filter(Tract_FIPS %in% c('36047051900','36081067100'))
# leaflet() %>%
#   addTiles() %>%
#   addPolygons(data = tract_final_spatial_tempproj, color = "blue", label = tract_final$Tract_FIPS,
#               popup = c(paste0('GEOID: ', tract_final$Tract_FIPS))) %>%
#   addPolygons(data = tract_final_spatial_tempproj %>% filter(Tract_FIPS %in% c('36047051900','36081067100')), color = "red", label = tract_final_spatial_tempproj$Tract_FIPS,
#               popup = c(paste0('GEOID: ', tract_final$Tract_FIPS)))

## CREATE: PART 1 - CLUSTER PROFILES
# Colors
# Explore color options
  # colorBlindness::displayAllColors(paletteMartin, color="white")
  # colors_colorblindsafe <- paletteMartin[c(5, 6, 9, 11, 14, 15)] %>% unname()
colors_colorblindsafe <- c('#ffff6d','#ffb6db','#24ff24','#6db6ff','#920000','#490092') 

# Alt: colors_colorblindsafe <- safeColors[2:7] %>% unname()
# Ref: https://www.r-graph-gallery.com/circular-barplot.html
nvi_cat_levels <- c('Demographics', 'Economic', 'Residential', 'Health Status')
# Data Prep
fig4b_part1_df <- tract_final %>% 
  dplyr::group_by(nvi_cluster) %>% 
  dplyr::summarize(n_obs = n(),
                   mean_nvi = mean(nvi),
                   q1_nvi = quantile(nvi, 0.25),
                   q3_nvi = quantile(nvi, 0.75),
                   iqr_nvi = paste0(round(q1_nvi, 2), "-", round(q3_nvi,2)),
                   mean_demo_score = mean(score_demo),
                   mean_economic_score = mean(score_economic),
                   mean_residential_score = mean(score_residential),
                   mean_healthstatus_score = mean(score_healthstatus)) %>% 
  tidyr::gather(nvi_cat_prelim, mean, mean_demo_score:mean_healthstatus_score) %>% 
  dplyr::mutate(nvi_cat = case_when(nvi_cat_prelim == 'mean_demo_score'         ~ nvi_cat_levels[1],
                                    nvi_cat_prelim == 'mean_economic_score'     ~ nvi_cat_levels[2],
                                    nvi_cat_prelim == 'mean_residential_score'  ~ nvi_cat_levels[3],
                                    nvi_cat_prelim == 'mean_healthstatus_score' ~ nvi_cat_levels[4]) %>% factor(nvi_cat_levels),
                nvi_cluster_label = paste0("Cluster ", nvi_cluster, " (n = ", n_obs, ",\nMean NVI = ", round(mean_nvi,2), ")"))
# Plotting
fig4b_part1_prelim <- ggplot(fig4b_part1_df, aes(x = nvi_cat, y = mean, fill = nvi_cluster_label, pattern = nvi_cat)) +
  ggpattern::geom_bar_pattern(stat = "identity", color = "black") +
  geom_text(aes(label = round(mean, 2), y = mean + 0.25), size = 4) + # y = mean + 0.5
  coord_polar(start = (3*pi)/2, direction = -1, clip = "off") +
  ylim(-0.2, 1.1) + #1.1
  scale_fill_manual(values = colors_colorblindsafe, labels = c("1","2","3","4","5","6")) +
  ggpattern::scale_pattern_manual(values = c("circle","stripe","crosshatch","none")) +
  facet_wrap(. ~ nvi_cluster_label, nrow = 2) +
  labs(fill = "Neighborhood\nVulnerability Index\nCluster", pattern = "Domain Score") +
  theme_minimal() +
  guides(fill = guide_legend(override.aes = list(pattern = "none"), nrow = 2, title.position = "top", title.hjust = 0.5, order = 1), # fill = FALSE 
         pattern = guide_legend(override.aes = list(fill = "white"), nrow = 6, title.position = "top", title.hjust = 0.5, order = 2)) +
  theme(legend.position = "right",
        legend.title = element_text(face = "bold"),
        legend.box = 'vertical',
        axis.title = element_blank(),
        axis.text = element_blank(),
        strip.text.x = element_text(size = 12, face = "bold"),
        panel.grid = element_blank(),
        plot.margin = unit(c(t = 0, r = 0, b = 0, l = 0), "cm"),
        panel.spacing = unit(0, "lines")) # strip.background.y = element_rect("transparent")
fig4b_part1_plot_build <- ggplot_build(fig4b_part1_prelim)
for(i in 1:6){
  fig4b_part1_plot_build[["layout"]][["panel_params"]][[i]][["r.range"]][2] <- 0.45
}
fig4b_part1 <- ggplot_gtable(fig4b_part1_plot_build)


## CREATE: PART 2 - CLUSTER MAP
gen_fig4b_part2 <- function(dfin, dfexcl, fill_var, fill_color, fill_label, fill_increment){
  plot_choro <- ggplot() + 
    geom_sf(data = dfin, aes(fill = !!sym(fill_var)), color = NA) + 
    geom_sf(data = dfexcl, color = "gray90", fill = "gray90") + 
    theme_minimal() + 
    scale_fill_manual(values = fill_color, guide = guide_legend(direction = "horizontal", title.position = "top")) +
    labs(fill = fill_label) + 
    # guides(fill = guide_legend(nrow = 1, title.position = "top", title.hjust = 0.5)) +
    theme(legend.position = "none", # legend.title = element_text(face = "bold", size = 12), legend.text = element_text(size = 9), legend.spacing.x = unit(0.5, 'cm'), 
          axis.line = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank(),
          panel.grid = element_blank(),
          plot.margin = unit(c(t = 0, r = 0, b = 0, l = 0), "cm")) 
  return(plot_choro)
}
fig4b_part2 <- gen_fig4b_part2(dfin = tract_final_spatial, dfexcl = tract_exclude_spatial, fill_var = "nvi_cluster", fill_color = colors_colorblindsafe, fill_label = "Neighborhood Vulnerability Index Cluster") 

## COMBINED PLOT
fig4b <- ggarrange(fig4b_part1, ggarrange(fig4b_part2, ggplot() + theme_void(), ncol = 2), nrow = 2) + theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0, "cm"))
fig4b
# EXPORT
ggsave('figs/fig4_partb.png', width = 7, height = 9) # pixels: 1200 x 900 width = 13, height = 12

# View average NVI domains scores by cluster for text
fig4b_part1_df %>% distinct(nvi_cluster, mean_nvi, iqr_nvi)
tract_final %>% count(nvi_cluster, nvi_cluster_orig)

```



## SUPPLEMENTAL FIGURE 1. Box plots
```{r}
# note changed colors for overall and residential
figs1_df <- tract_final %>% 
  dplyr::select(Tract_FIPS, county, nvi, score_demo, score_economic, score_residential, score_healthstatus) %>% 
  tidyr::gather(key = index_type, value = index, nvi, score_demo, score_economic, score_residential, score_healthstatus) %>% 
  dplyr::mutate(index_type = case_when(index_type == "nvi" ~ "Overall:\nNeighborhood\nVulnerability Index",
                                       index_type == "score_demo" ~ "Domain:\nDemographic\nScore",
                                       index_type == "score_economic" ~ "Domain:\nEconomic\nScore",
                                       index_type == "score_residential" ~ "Domain:\nResidential\nScore",
                                       index_type == "score_healthstatus" ~ "Domain:\nHealth Status\nScore") %>% 
                                       factor(levels = c("Overall:\nNeighborhood\nVulnerability Index","Domain:\nDemographic\nScore","Domain:\nEconomic\nScore","Domain:\nResidential\nScore","Domain:\nHealth Status\nScore",
                                                         'Neighborhood Deprivation Index','Social Vulnerability Index')),
                part = 'NVI') %>% 
  rbind(ndi_clean %>% dplyr::mutate(index_type = 'Neighborhood Deprivation Index', part = 'NDI') %>% rename(index = ndi),
        svi_clean %>% dplyr::mutate(index_type = 'Social Vulnerability Index', part = 'SVI') %>% rename(index = svi)) 
### CREATE: Plots
  # NVI
figs1a <- ggplot(figs1_df %>% filter(part == 'NVI'), aes(y = county, x = index, fill = index_type)) +
  geom_boxplot() +
  facet_grid(index_type ~ ., scale = "free") +
  # labs(x = "Index", y = "County") +
  labs(title = "Neighborhood Vulnerability Index") +
  scale_fill_manual(values = c("#D55E00","#0072B2","#009E73","#E69F00","#CC79A7")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_blank(),  axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), strip.text.y = element_text(size = 20), plot.title = element_text(face = "bold", hjust = 0.5, size = 20))
  # NDI
figs1b <- ggplot(figs1_df %>% filter(part == 'NDI'), aes(y = county, x = index, fill = index_type)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#D55E00")) +
  labs(title = "Neighborhood Deprivation Index") +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_blank(),  axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), plot.title = element_text(face = "bold", hjust = 0.5, size = 20))
  # SVI
figs1c <- ggplot(figs1_df %>% filter(part == 'SVI'), aes(y = county, x = index, fill = index_type)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#D55E00")) +
  labs(title = "Social Vulnerability Index") +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_blank(),  axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), plot.title = element_text(face = "bold", hjust = 0.5, size = 20))
### EXPORT: Plots
# fig3_labs <- c("A","B","C") # c("Neighborhood Vulnerability Index","Neighborhood Deprivation Index","Social Vulnerability Index")
figs1 <- ggarrange(figs1a, figs1b, figs1c, nrow = 3, font.label = list(size = 25), align = "v", heights = c(4,1,1)) #labels = fig3_labs,
annotate_figure(figs1,left = text_grob("County", rot = 90, hjust = 1, size = 20), bottom = text_grob("Index", size = 20))
ggsave('figs/figs1.png', width = 15, height = 16)
```

### SUPPLEMENTAL FIGURE 3 Sensitivity analysis: Race/ethnicity
```{r}
### Clean data 
# CREATE: Function
clean_toxpi_sa <- function(datain, datacounty){
  dataout <- datain %>% 
    dplyr::rename(Tract_FIPS = Source, nvi = `ToxPi Score`) %>% 
    dplyr::mutate(Tract_FIPS = as.character(Tract_FIPS) %>% trimws()) %>% 
    dplyr::select(-c(`HClust Group`, `KMeans Group`, Name)) %>% 
    # Demographics
    dplyr::rename_at(vars(starts_with('DemographicsAge')), funs(paste0('score_demo_age'))) %>% 
    dplyr::rename_at(vars(starts_with('DemographicsFemaleLed')), funs(paste0('score_demo_femaleled'))) %>% 
    dplyr::rename_at(vars(starts_with('DemographicsRaceEth!')), funs(paste0('score_demo_raceeth'))) %>% 
    dplyr::rename_at(vars(starts_with('DemographicsImmigration')), funs(paste0('score_demo_immigration'))) %>% 
    dplyr::rename_at(vars(starts_with('DemographicsDisability')), funs(paste0('score_demo_disability'))) %>% 
    dplyr::rename_at(vars(starts_with('DemographicsSingleParent')), funs(paste0('score_demo_singleparent'))) %>% 
    dplyr::rename_at(vars(starts_with('DemographicsMobility')), funs(paste0('score_demo_mobility'))) %>% 
    dplyr::rename_at(vars(starts_with('DemographicsLiveAlone')), funs(paste0('score_demo_livealone'))) %>% 
    # Economic
    dplyr::rename_at(vars(starts_with('Economic IndicatorsIncomePoverty')), funs(paste0('score_economic_incomepoverty'))) %>% 
    dplyr::rename_at(vars(starts_with('Economic IndicatorsServiceManualJobs')), funs(paste0('score_economic_servicemanual'))) %>% 
    dplyr::rename_at(vars(starts_with('Economic IndicatorsGini')), funs(paste0('score_economic_gini'))) %>% 
    dplyr::rename_at(vars(starts_with('Economic IndicatorsEmployment')), funs(paste0('score_economic_employment'))) %>% 
    dplyr::rename_at(vars(starts_with('Economic IndicatorsEducation')), funs(paste0('score_economic_education'))) %>% 
    dplyr::rename_at(vars(starts_with('Economic IndicatorsVehicleAvail')), funs(paste0('score_economic_vehicleavail'))) %>% 
    # Residential
    dplyr::rename_at(vars(starts_with('Residential DensityPopDensity')), funs(paste0('score_residential_popdensity'))) %>% 
    dplyr::rename_at(vars(starts_with('Residential DensityGroupQuarters')), funs(paste0('score_residential_groupquarters'))) %>% 
    dplyr::rename_at(vars(starts_with('Residential DensityOccPerRoom')), funs(paste0('score_residential_occperroom'))) %>% 
    dplyr::rename_at(vars(starts_with('Residential DensityStrAge')), funs(paste0('score_residential_structage'))) %>% 
    dplyr::rename_at(vars(starts_with('Residential DensityStrAttached')), funs(paste0('score_residential_structattach'))) %>% 
    dplyr::rename_at(vars(starts_with('Residential DensityMove1Yr')), funs(paste0('score_residential_move1yr'))) %>% 
    dplyr::rename_at(vars(starts_with('Residential DensityVacancy')), funs(paste0('score_residential_vacancy'))) %>% 
    # Health status
    dplyr::rename_at(vars(starts_with('Health StatusLifestyle')), funs(paste0('score_healthstatus_lifestyle'))) %>% 
    dplyr::rename_at(vars(starts_with('Health StatusCondition')), funs(paste0('score_healthstatus_condition'))) %>% 
    dplyr::rename_at(vars(starts_with('Health StatusPreventive')), funs(paste0('score_healthstatus_preventive'))) %>% 
    dplyr::rename_at(vars(starts_with('Health StatusLackInsurance')), funs(paste0('score_healthstatus_lackinsurance'))) %>% 
    dplyr::mutate(score_economic = (score_economic_incomepoverty + score_economic_servicemanual + score_economic_gini + score_economic_employment + score_economic_education + score_economic_vehicleavail)/6,
                  score_residential = (score_residential_popdensity + score_residential_groupquarters + score_residential_occperroom + score_residential_structage + score_residential_structattach + score_residential_move1yr + score_residential_vacancy)/7,
                  score_healthstatus = (score_healthstatus_lifestyle + score_healthstatus_condition + score_healthstatus_preventive + score_healthstatus_lackinsurance)/4) %>% 
    dplyr::left_join(datacounty, by = "Tract_FIPS")
  return(dataout)
}
# DATA CLEANING
  # Race/eth: no overlapping
tract_toxpi_raceeth_no_olap_df <- clean_toxpi_sa(tract_toxpi_pre_raceeth_no_olap, tract_nyc_county) %>% 
  dplyr::mutate(score_demo = (score_demo_age + score_demo_femaleled + score_demo_raceeth + score_demo_immigration + score_demo_disability + score_demo_singleparent + score_demo_mobility + score_demo_livealone)/8) 
  # Race/eth: overlapping
tract_toxpi_raceeth_olap_df <- clean_toxpi_sa(tract_toxpi_pre_raceeth_olap, tract_nyc_county) %>% 
  dplyr::mutate(score_demo = (score_demo_age + score_demo_femaleled + score_demo_raceeth + score_demo_immigration + score_demo_disability + score_demo_singleparent + score_demo_mobility + score_demo_livealone)/8) 
  # Race/eth + Hisp
tract_toxpi_raceeth_hsub_df <- clean_toxpi_sa(tract_toxpi_pre_raceeth_hsub, tract_nyc_county) %>% 
  dplyr::rename_at(vars(starts_with('DemographicsRaceEthHisp')), funs(paste0('score_demo_raceeth_hisp'))) %>% 
  dplyr::mutate(score_demo = (score_demo_age + score_demo_femaleled + score_demo_raceeth + score_demo_raceeth_hisp + score_demo_immigration + score_demo_disability + score_demo_singleparent + score_demo_mobility + score_demo_livealone)/9)

##### PLOTS
### BOXPLOT
# PREP data
figs3_df <- rbind(tract_final %>% dplyr::select(nvi, score_demo, county) %>% dplyr::mutate(index_type = "Main"),
               tract_toxpi_raceeth_no_olap_df %>% dplyr::select(nvi, score_demo, county) %>% dplyr::mutate(index_type = "Non-Overlapping Race/Ethnicity\nFeatures"),
               tract_toxpi_raceeth_olap_df %>% dplyr::select(nvi, score_demo, county) %>% dplyr::mutate(index_type = "Overlapping Race/Ethnicity\nCategories"),
               tract_toxpi_raceeth_hsub_df %>% dplyr::select(nvi, score_demo, county) %>% dplyr::mutate(index_type = "Hispanic Separate")) %>% 
  dplyr::mutate(all = "All Counties", index_type = factor(index_type, levels = c("Main",
                                                           "Non-Overlapping Race/Ethnicity\nFeatures",
                                                           "Overlapping Race/Ethnicity\nCategories",
                                                           "Hispanic Separate")))
figs3_sub_df <- rbind(tract_toxpi_raceeth_no_olap_df %>% dplyr::select(nvi, score_demo_raceeth, county) %>% dplyr::mutate(index_type = "Non-Overlapping Race/Ethnicity\nFeatures"),
                      tract_toxpi_raceeth_olap_df %>% dplyr::select(nvi, score_demo_raceeth, county) %>% dplyr::mutate(index_type = "Overlapping Race/Ethnicity\nCategories\nRace/Ethnicity Subdomain"),
                      tract_toxpi_raceeth_hsub_df %>% dplyr::select(nvi, score_demo_raceeth, county) %>% dplyr::mutate(index_type = "Hispanic Separate\nRace Subdomain"),
                      tract_toxpi_raceeth_hsub_df %>% dplyr::select(nvi, score_demo_raceeth_hisp, county) %>% dplyr::rename(score_demo_raceeth = score_demo_raceeth_hisp) %>% dplyr::mutate(index_type = "Hispanic Separate\nHispanic Subdomain")) %>% 
  dplyr::mutate(all = "All Counties", index_type = factor(index_type, levels = c("Non-Overlapping Race/Ethnicity\nFeatures",
                                                           "Overlapping Race/Ethnicity\nCategories\nRace/Ethnicity Subdomain",
                                                           "Hispanic Separate\nRace Subdomain",
                                                           "Hispanic Separate\nHispanic Subdomain")))


# CREATE: Plots
  # NVI
figs3a <- ggplot(figs3_df, aes(x = nvi, fill = index_type)) +
  geom_boxplot(aes(y = county)) +
  geom_boxplot(aes(y = all)) +
  scale_y_discrete(limits = rev) +
  geom_hline(yintercept = 5.5, color = "gray20", linetype = "dashed") +
  labs(title = "Overall NVI", fill = "NVI:\nRace/Ethnicity Categorization") +
  scale_fill_manual(values = c("#D55E00", "gray40", "gray60", "gray80")) + # limits = rev
  guides(fill = guide_legend(reverse = TRUE)) +
  theme_bw() +
  theme(axis.title.x = element_blank(),  axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), panel.grid.major.y = element_blank(), strip.text.y = element_text(size = 20), plot.title = element_text(face = "bold", hjust = 0.5, size = 20), legend.title = element_text(size = 16), legend.text = element_text(size = 14))
  # Domain: Demographic 
figs3b <- ggplot(figs3_df, aes(x = score_demo, fill = index_type)) +
  geom_boxplot(aes(y = county)) +
  geom_boxplot(aes(y = all)) +
  scale_y_discrete(limits = rev) +
  geom_hline(yintercept = 5.5, color = "gray20", linetype = "dashed") +
  labs(title = "Domain: Demographic Score", fill = "Demographic Domain:\nRace/Ethnicity Categorization") +
  scale_fill_manual(values = c("#0072B2", "gray40", "gray60", "gray80")) +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme_bw() +
  theme(axis.title.x = element_blank(),  axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), panel.grid.major.y = element_blank(), plot.title = element_text(face = "bold", hjust = 0.5, size = 20), legend.title = element_text(size = 16), legend.text = element_text(size = 14))
  # Subdomain: Race/ethnicity
figs3c <- ggplot(figs3_sub_df, aes(x = score_demo_raceeth, fill = index_type)) +
  geom_boxplot(aes(y = county)) +
  geom_boxplot(aes(y = all)) +
  scale_y_discrete(limits = rev) +
  geom_hline(yintercept = 5.5, color = "gray20", linetype = "dashed") +
  labs(title = "Subdomain: Race/Ethnicity Score", fill = "Race/Ethnicity Subdomain:\nRace/Ethnicity Categorization") +
  scale_fill_manual(values = c("gray40", "gray60", "gray80", "gray80")) +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme_bw() +
  theme(axis.title.x = element_blank(),  axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), panel.grid.major.y = element_blank(), plot.title = element_text(face = "bold", hjust = 0.5, size = 20), legend.title = element_text(size = 16), legend.text = element_text(size = 14))
### MERGE AND EXPORT: Plots
figs3 <- ggarrange(figs3a, figs3b, figs3c, nrow = 3, font.label = list(size = 25), align = "v", heights = c(1,1,1)) 
annotate_figure(figs3,left = text_grob("County", rot = 90, hjust = 1, size = 20), bottom = text_grob("Index", size = 20))
ggsave('figs/figs3.png', width = 15, height = 20)

```


### Descriptives (for text)
```{r}
# Prep data
text_df <- tract_final %>% 
  dplyr::left_join(ndi_clean, by = "Tract_FIPS") %>% 
  dplyr::left_join(svi_clean, by = "Tract_FIPS")


# Function to generate summary stats
gen_summary <- function(dfin, var, var_label){
  var <- sym(var)
  overall <- dfin %>% 
    dplyr::summarize(mean = mean(!!var), sd = sd(!!var), median = median(!!var), min = min(!!var), max = max(!!var), q1 = quantile(!!var, probs = 0.25), q3 = quantile(!!var, probs = 0.75), iqr = paste0(q1 %>% round(2) %>% format(nsmall = 2),"-",q3 %>% round(2) %>% format(nsmall = 2)), median_iqr = paste0(median %>% round(2) %>% format(nsmall = 2)," (IQR = ",iqr,")"), median_iqr2 = paste0(median %>% round(2) %>% format(nsmall = 2),", IQR = ",iqr), median_iqr3 = paste0("median = ", median_iqr2)) %>% 
    dplyr::mutate(county = "Overall")
  bycounty <- dfin %>% 
    dplyr::group_by(county) %>% 
    dplyr::summarize(mean = mean(!!var), sd = sd(!!var), median = median(!!var), min = min(!!var), max = max(!!var), q1 = quantile(!!var, probs = 0.25), q3 = quantile(!!var, probs = 0.75), iqr = paste0(q1 %>% round(2) %>% format(nsmall = 2),"-",q3 %>% round(2) %>% format(nsmall = 2)), median_iqr = paste0(median %>% round(2) %>% format(nsmall = 2)," (IQR = ",iqr,")"), median_iqr2 = paste0(median %>% round(2) %>% format(nsmall = 2),", IQR = ",iqr), median_iqr3 = paste0("median = ", median_iqr2)) %>% 
    dplyr::mutate(county = paste0("   ", county))
  combined <- rbind(overall, bycounty) %>% 
    dplyr::mutate(score = var_label) %>% 
    dplyr::relocate(score, county, mean, sd, median, iqr, min, max, median_iqr, median_iqr2, median_iqr3, .before = mean)
  return(combined)
}

# Generate summary stats
text_summary <- rbind(gen_summary(text_df, "nvi", "Neighborhood Vulnerability Index (NVI)")
                     ,gen_summary(text_df, "ndi", "Neighborhood Deprivation Index (NDI)")
                     ,gen_summary(text_df, "svi", "Social Vulnerability Index (SVI)")
                     ,gen_summary(text_df, "score_demo", "NVI Domain: Demographics")
                     ,gen_summary(text_df, "score_economic", "NVI Domain: Economic Indicators")
                     ,gen_summary(text_df, "score_residential", "NVI Domain: Residential Characteristics")
                     ,gen_summary(text_df, "score_healthstatus", "NVI Domain: Health Status"))
text_summary %>% head()

# text_summary %>% View()
export(text_summary, 'text_summary.xlsx')

# OVERALL CORRELATIONS
cor.test(text_df$nvi, text_df$ndi, method = "spearman")
cor.test(text_df$nvi, text_df$svi, method = "spearman")


```

