---
title: "Neighborhood Environmental Vulnerability Index, 2019: Create Figures and Tables"
author: 
- "Stephen P. Uong" 
- "Contributors: Jiayi Zhou, Jeanette A. Stingone"
date: "10/28/2021"
output: 
  html_document: 
    toc: true
    toc_float: true
---

Below are steps to create figures and tables for the NEVI.

In this document, we have included code we used to generate results and figures that we used when evaluating the NEVI and comparing the NEVI to the Neighborhood Deprivation Index (NDI)/ Social Vulnerability Index (SVI).

### 1. Setup
Set the working directory to one folder up from the RMarkdown file for later data export.
```{r, setup}
knitr::opts_knit$set(root.dir = '..') 
```

### 2. Load Required Libraries

Load the following required libraries.
```{r, libraries}
library(tidyverse)
library(rio)
library(skimr)
library(janitor)
library(cluster)
library(factoextra)
# library(Amelia)
# library(janitor)
# library(cluster)
# Mapping
library(sf)
library(ggpubr)
library(ggsn)
library(ggspatial)
library(nycgeo)
library(tigris)
# https://yutannihilation.github.io/ggsflabel/index.html
library(ggsflabel) # devtools::install_github("yutannihilation/ggsflabel")
# Plots
library(ggpubr)
library(colorBlindness)
library(ggpattern) # Ref: https://coolbutuseless.github.io/package/ggpattern/index.html
library(gplots) # heat map
```


### 3: Importing the Index Generated from the ToxPi GUI, NEVI features, NDI, and SVI
```{r}
setwd('C:/Users/steph/OneDrive - cumc.columbia.edu/Phi/My Courses/Columbia/Work/Research - Jeanette/projects/covid-vulnerability-index/comms/git nvi')

### IMPORT: Cleaned NEVI data with the NEVI, NEVI features, and NEVI clusters
tract_final <- readRDS(file = "data/processed/nevi_final.rds")
### IMPORT: NYC specific Census tract boundaries
tract_nyc_spatial <- nycgeo::nyc_boundaries(geography = "tract") %>% 
  dplyr::transmute(Tract_FIPS = geoid) %>% 
  sf::st_transform(crs = 2263)
tract_nyc_borough <- nycgeo::nyc_boundaries(geography = "tract") %>% 
  dplyr::as_tibble() %>% 
  dplyr::select(-geometry) %>% 
  dplyr::transmute(Tract_FIPS = geoid,
                   borough = borough_name)
borough_nyc_spatial <- nycgeo::nyc_boundaries(geography = "borough") %>%
  dplyr::arrange(borough_name) %>% 
  sf::st_transform(crs = 2263)

### IMPORT: Toxpi data before cleaning (for 8 clusters)
tract_toxpi_pre <- import('data/processed/toxpi/nevi_toxpi_results.csv')

### IMPORT: Exclusions List
tract_exclusions_list <- import("data/processed/EXCLUSION_LIST_20210628.csv")
tract_exclusions_list_id <- tract_exclusions_list %>% 
  dplyr::filter(flag_exclude_FINAL == 0) %>% 
  dplyr::transmute(SID = as.character(GEOID))

### IMPORT: NDI, SVI
ndi <- import("data/processed/NDI.NYC.2086t.csv")
svi <- import("data/raw/svi.csv")

ndi_clean <- ndi %>% dplyr::mutate(Tract_FIPS = GEOID %>% as.character() %>% str_pad(11,side="left",pad="0")) %>% dplyr::rename(ndi = NDI_score) %>% dplyr::select(Tract_FIPS, ndi) %>% 
  dplyr::left_join(tract_nyc_borough, by = "Tract_FIPS")
svi_clean <- svi %>% dplyr::mutate(Tract_FIPS = FIPS %>% as.character() %>% str_pad(11,side="left",pad="0")) %>% dplyr::rename(svi = RPL_THEMES) %>% dplyr::select(Tract_FIPS, svi) %>% 
  dplyr::filter(svi != -999) %>% 
  dplyr::left_join(tract_nyc_borough, by = "Tract_FIPS") %>% 
  dplyr::inner_join(tract_exclusions_list_id %>% dplyr::transmute(Tract_FIPS = SID), by = "Tract_FIPS")

## IMPORT: Race/ethnicity sensitivity analysis
tract_toxpi_pre_raceeth_no_olap <- import('data/processed/toxpi/nevi_toxpi_raceeth_no_olap_results.csv')
tract_toxpi_pre_raceeth_olap <- import('data/processed/toxpi/nevi_toxpi_raceeth_olap_results.csv')
tract_toxpi_pre_raceeth_hsub <- import('data/processed/toxpi/nevi_toxpi_raceeth_hsub_results.csv')
```

### Create spatial objects
```{r}
# Spatial: crs = "+init=epsg:2263"
  # Included
tract_final_spatial <- tract_nyc_spatial %>% 
  dplyr::inner_join(tract_final %>% dplyr::mutate(Tract_FIPS = as.character(Tract_FIPS) %>% trimws()), by = "Tract_FIPS") # Checked: did not exclude any tracts, 8/11
  # Excluded
tract_exclude_spatial <- tract_nyc_spatial %>% 
  dplyr::anti_join(tract_final %>% dplyr::mutate(Tract_FIPS = as.character(Tract_FIPS) %>% trimws()), by = "Tract_FIPS") 
```


# FIGURES
## FIGURE 1: Domains and subdomains
```{r mapping, include=FALSE}
#### CREATE: Functions to create maps
gen_fig1_main <- function(dfin, dfexcl, dfborough, fill_var, fill_color, fill_label, lab_borough){
  plot_choro <- ggplot() + 
    # tracts
    geom_sf(data = dfin, aes(fill = !!sym(fill_var)), color = NA) + 
    # tracts, excluded
    geom_sf(data = dfexcl, color = "gray30", fill = "gray30") + 
    theme_minimal() + # Alt: ggthemes::theme_map()
        # boroughs
    geom_sf(data = dfborough, color = 'gray20', fill = NA, size = 0.6) +
    geom_sf_label_repel(data = dfborough, aes(label = !!sym(lab_borough)), force = 0, nudge_x = c(40000, -45000, -30000, 30000, 35000), nudge_y = c(20000, 20000, 10000, -50000, -20000), seed = 1, size = 6) +
    ggsn::scalebar(data = dfborough, dist = 10, dist_unit = "mi", height = 0.01, st.size = 5, st.dist = 0.05, transform = FALSE, model = 'WGS84', location = "bottomright") + # scale bar
    ggsn::north(data = dfborough, location = "topleft", symbol = 10, scale = 0.10) + # north arrow, symbol = 16, scale = 0.17
        # Alt: ggspatial::annotation_north_arrow(pad_x = unit(2.5, "cm"), pad_y = unit(-1.75, "cm")) +
        # Alt: ggspatial::annotation_scale(plot_unit = "mi", pad_x = unit(7.5, "cm"), pad_y = unit(-1, "cm")) #bgcolor("lightcyan") + #snow2 
    labs(fill = fill_label) + 
    scale_fill_gradient(low = "whitesmoke", high = fill_color) + # na.value = "gray30"
    theme(legend.title = element_text(face = "bold", size = 15), # plot.title = element_text(hjust = 0.75, face = "bold", size = 15),
          legend.text = element_text(size = 12),
          legend.spacing.x = unit(0.5, 'cm'),
          legend.position = "right",
          axis.line = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank(),
          panel.grid = element_blank()) # plot.caption = element_text(hjust = 0, size = 10, face = "italic"), legend.position = c(1,0.5), plot.background = element_rect(fill = "lightcyan")
  return(plot_choro)
}
gen_fig1_part <- function(dfin, dfexcl, dfborough, fill_var, fill_color, fill_label, fill_label_just, fill_increment){
  plot_choro <- ggplot() + 
    geom_sf(data = dfin, aes(fill = !!sym(fill_var)), color = NA) + 
    geom_sf(data = dfexcl, color = "gray30", fill = "gray30") + 
    geom_sf(data = dfborough, color = 'gray20', fill = NA, size = 0.6) +
    theme_minimal() + 
    scale_fill_gradient(low = "whitesmoke", high = fill_color, breaks = seq(0, 1, fill_increment)) + 
    labs(fill = fill_label) + 
    theme(legend.title = element_text(face = "bold", size = 14, vjust = fill_label_just),  # vertically center legend with its label - does not work well
          legend.text = element_text(size = 10),
          legend.spacing.x = unit(0.5, 'cm'),
          legend.position = "right",
          axis.line = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank(),
          panel.grid = element_blank())
  return(plot_choro)
}

### CREATE: Maps
  # Overall Index
fig1_toxpi <- gen_fig1_main(dfin = tract_final_spatial, dfexcl = tract_exclude_spatial, dfborough = borough_nyc_spatial, fill_var = "nevi", fill_color = '#954100',
fill_label = "Neighborhood\nEnvironmental\nVulnerability Index", lab_borough = 'borough_name')
  # Domain Scores
fig1_demo <- gen_fig1_part(dfin = tract_final_spatial, dfexcl = tract_exclude_spatial, dfborough = borough_nyc_spatial, fill_var = "score_demo", fill_color = '#004F7C', fill_label = "Demographics Score", fill_label_just = 0.5, fill_increment = 0.1)
fig1_econ <- gen_fig1_part(dfin = tract_final_spatial, dfexcl = tract_exclude_spatial, dfborough = borough_nyc_spatial,fill_var = "score_economic", fill_color = '#006E50', fill_label = "Economic\nIndicators Score", fill_label_just = 0.5, fill_increment = 0.2)
fig1_resd <- gen_fig1_part(dfin = tract_final_spatial, dfexcl = tract_exclude_spatial, dfborough = borough_nyc_spatial,fill_var = "score_residential", fill_color = '#A16F00', fill_label = "Residential Density\nand Characteristics\nScore", fill_label_just = 4, fill_increment = 0.2)
fig1_hlth <- gen_fig1_part(dfin = tract_final_spatial, dfexcl = tract_exclude_spatial, dfborough = borough_nyc_spatial,fill_var = "score_healthstatus", fill_color = '#8E5474', fill_label = "Health Behavior,\nOutcomes,\nPrevention Practices,\nand Access Score", fill_label_just = 4, fill_increment = 0.2)
## MERGE: Maps
  # Alt: gridExtra::grid.arrange(fig1_toxpi, fig1_demo, fig1_econ, fig1_resd, fig1_hlth, ncol = 4, nrow = 2, layout_matrix = rbind(c(1,1,2,3), c(1,1,4,5))) 
  # Ref: https://www.datanovia.com/en/lessons/combine-multiple-ggplots-into-a-figure/
fig1_part1 <- ggarrange(fig1_toxpi, fig1_demo, nrow = 2, labels = c("A","B"), align = "v", font.label = list(size = 25)) + theme(plot.margin = margin(t = 0, r = 0, b = 1, l = 0, "cm"))
fig1_part2 <- ggarrange(fig1_econ, fig1_resd, nrow = 2, labels = c("C","D"), align = "v", font.label = list(size = 25)) + theme(plot.margin = margin(t = 0, r = 0, b = 1, l = 0, "cm"))
fig1_part3 <- ggarrange(fig1_hlth, labels = "E", font.label = list(size = 25)) + theme(plot.margin = margin(t = 0, r = 0, b = 1, l = 0, "cm"))

## EXPORT
fig1_part1
ggsave('figs/fig1_part1.png', width = 10, height = 15) # pixels: 1200 x 900, width = 15, height = 12
fig1_part2
ggsave('figs/fig1_part2.png', width = 10, height = 15) 
fig1_part3
ggsave('figs/fig1_part3.png', width = 10, height = 8) 

```

## FIGURE 2. MAP: NEVI, NDI, SVI
```{r}
#### CREATE: Functions to create maps
gen_fig2_main <- function(dfin, dfexcl, dfborough, fill_var, fill_color, fill_label){
  plot_choro <- ggplot() + 
    geom_sf(data = dfin, aes(fill = !!sym(fill_var)), color = NA) + 
    geom_sf(data = dfexcl, color = "gray30", fill = "gray30") + 
    geom_sf(data = dfborough, color = 'gray20', fill = NA, size = 0.6) +
    theme_minimal() + # Alt: ggthemes::theme_map()
    ggsn::scalebar(data = dfin, dist = 20, dist_unit = "mi", height = 0.01, st.size = 5, st.dist = 0.05, transform = FALSE, model = 'WGS84', location = "topleft") + # scale bar, bottomright
    ggsn::north(data = dfin, location = "bottomright", symbol = 10, scale = 0.10) + # north arrow, symbol = 16, scale = 0.17, topleft
        # Alt: ggspatial::annotation_north_arrow(pad_x = unit(2.5, "cm"), pad_y = unit(-1.75, "cm")) +
        # Alt: ggspatial::annotation_scale(plot_unit = "mi", pad_x = unit(7.5, "cm"), pad_y = unit(-1, "cm")) #bgcolor("lightcyan") + #snow2 
    labs(fill = fill_label) + 
    #scale_fill_gradient(low = "whitesmoke", high = fill_color) + # na.value = "gray30"
    scale_fill_manual(values = fill_color, guide = guide_legend(direction = "horizontal", title.position = "top")) + # cannot seem to get title to center: title.hjust = 0.5
    theme(legend.title = element_text(face = "bold", size = 15), # plot.title = element_text(hjust = 0.75, face = "bold", size = 15),
          legend.text = element_text(size = 12),
          legend.spacing.x = unit(0.5, 'cm'),
          legend.position = "bottom",
          legend.box = "vertical",
          axis.line = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank(),
          panel.grid = element_blank())  # plot.caption = element_text(hjust = 0, size = 10, face = "italic"), legend.position = c(1,0.5), plot.background = element_rect(fill = "lightcyan")
    #guides(size = guide_legend(title.position = "top", title.hjust = 0.5))
  return(plot_choro)
}
gen_fig2_part <- function(dfin, dfexcl, dfborough, fill_var, fill_color, fill_label, fill_label_just, fill_increment){
  plot_choro <- ggplot() + 
    geom_sf(data = dfin, aes(fill = !!sym(fill_var)), color = NA) + 
    geom_sf(data = dfexcl, color = "gray30", fill = "gray30") + 
    geom_sf(data = dfborough, color = 'gray20', fill = NA, size = 0.6) +
    theme_minimal() + 
    scale_fill_manual(values = fill_color, guide = guide_legend(direction = "horizontal", title.position = "top")) +
    labs(fill = fill_label) + 
    theme(legend.title = element_text(face = "bold", size = 14, vjust = fill_label_just),  # vertically center legend with its label - does not work well
          legend.text = element_text(size = 10),
          legend.spacing.x = unit(0.5, 'cm'),
          legend.position = "bottom",
          axis.line = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank(),
          panel.grid = element_blank()) 
  return(plot_choro)
}

### PREP DATA
fig2_sf <- tract_final_spatial %>% 
  dplyr::select(Tract_FIPS, nevi) %>% 
  dplyr::left_join(ndi_clean, by = "Tract_FIPS") %>% 
  dplyr::left_join(svi_clean, by = "Tract_FIPS") %>% 
  dplyr::mutate(nevi_q = ntile(nevi, 4) %>% as.factor(),
                ndi_q = ntile(ndi, 4) %>% as.factor(),
                svi_q = ntile(svi, 4) %>% as.factor()) 

### CREATE: Maps for NEVI, NDI, SVI
red_fill <- c('#F6DECC', '#E59E66', '#d55e00', '#954100') # "#fee5d9","#fcae91","#fb6a4a","#cb181d"
fig2_toxpi <- gen_fig2_main(dfin = fig2_sf, dfexcl = tract_exclude_spatial, dfborough = borough_nyc_spatial, fill_var = "nevi_q", fill_color = red_fill, fill_label = "Neighborhood Environmental\nVulnerability Index Quartile") 
fig2_ndi <- gen_fig2_part(dfin = fig2_sf, dfexcl = tract_exclude_spatial, dfborough = borough_nyc_spatial, fill_var = "ndi_q", fill_color =  red_fill, fill_label = "Neighborhood Deprivation Index\nQuartile", fill_label_just = 0.5, fill_increment = 0.1) 
fig2_svi <- gen_fig2_part(dfin = fig2_sf, dfexcl = tract_exclude_spatial, dfborough = borough_nyc_spatial, fill_var = "svi_q", fill_color =  red_fill, fill_label = "Social Vulnerability Index\nQuartile", fill_label_just = 0.5, fill_increment = 0.2) 
## MERGE: Maps
fig2 <- ggarrange(fig2_toxpi, fig2_ndi, fig2_svi, ncol = 3, labels = c("A","B","C")) + theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0, "cm"))
# fig2 <- ggarrange(fig2_toxpi, ggarrange(fig2_ndi, fig2_svi, nrow = 2, labels = "B", font.label = list(size = 25)), ncol = 2, labels = "A", font.label = list(size = 25)) + theme(plot.margin = margin(t = 0, r = 0, b = 2, l = 0, "cm"))
fig2
ggsave('figs/fig2.png', width = 13, height = 6) # pixels: 1200 x 900 width = 13, height = 12

```

## FIGURE 3 / SUPPLEMENTAL FIGURE 2. Scatter plots
```{r}
### CREATE: Function to generate scatterplots
  # Ref, stat_cor: https://www.rdocumentation.org/packages/ggpubr/versions/0.4.0/topics/stat_cor
gen_fig3_scatterplot <- function(datain, var, var_label, var_breaks){
  plot_scatter <- ggplot(datain, aes(x = nevi, y = !!sym(var), color = nevi_domain)) +
    geom_point(alpha = 0.5, size = 0.8) + #  geom_smooth(method = lm) +
    ggpubr::stat_cor(aes(label = ..r.label..), method = "spearman", label.x.npc = 0.6, label.y.npc = 0.25, p.accuracy = 0.001, size = 6, color = "black") + # "~`,`~" # label.x.npc = 0.7, size = 4
    # ggpubr::stat_cor(aes(label = ..p.label..), method = "spearman", label.x.npc = 0.7, label.y.npc = 0.1, p.accuracy = 0.001, size = 4, color = "black") + # "~`,`~"
    facet_grid(borough ~ nevi_domain) + 
    scale_color_manual(values = c("#D55E00","#0072B2","#009E73","#E69F00","#CC79A7")) +
    scale_x_continuous(breaks = seq(-1, 0.8, 0.2), minor_breaks = seq(-1, 0.9, 0.1)) +
    scale_y_continuous(breaks = seq(-2, 6, var_breaks), minor_breaks = seq(-2, 6, var_breaks/2)) +
    theme_bw() +
    labs(x = "Neighborhood Environmental Vulnerability Index", y = var_label) +
    theme(legend.position = "none",
          axis.title.x = element_text(size = 20, margin = margin(t = 10)),
          axis.title.y = element_text(size = 20, margin = margin(r = 10)),
          axis.text.x = element_text(size = 13),
          axis.text.y = element_text(size = 13),
          strip.text.x = element_text(size = 14),
          strip.text.y = element_text(size = 14))
  return(plot_scatter)
}

### DATA PREP
fig3_nevi_df <- tract_final %>% 
  dplyr::select(Tract_FIPS, borough, nevi, score_demo, score_economic, score_residential, score_healthstatus) %>% 
  tidyr::gather(key = nevi_domain, value = nevi, nevi, score_demo, score_economic, score_residential, score_healthstatus) %>% 
  dplyr::mutate(nevi_domain = case_when(nevi_domain == "nevi" ~ "Overall:\nNeighborhood Environmental\nVulnerability Index",
                                       nevi_domain == "score_demo" ~ "Domain:\nDemographic Score",
                                       nevi_domain == "score_economic" ~ "Domain:\nEconomic Score",
                                       nevi_domain == "score_residential" ~ "Domain:\nResidential Score",
                                       nevi_domain == "score_healthstatus" ~ "Domain:\nHealth Status Score") %>% 
                                       factor(levels = c("Overall:\nNeighborhood Environmental\nVulnerability Index","Domain:\nDemographic Score","Domain:\nEconomic Score","Domain:\nResidential Score","Domain:\nHealth Status Score")))
### MERGE with NDI/SVI data
fig3_df <- fig3_nevi_df %>%
  dplyr::mutate(Tract_FIPS = as.character(Tract_FIPS)) %>% 
  dplyr::left_join(ndi_clean %>% dplyr::select(-borough), by = "Tract_FIPS") %>% 
  dplyr::left_join(svi_clean %>% dplyr::select(-borough), by = "Tract_FIPS")
### CREATE: Plots
fig3_ndi <- gen_fig3_scatterplot(datain = fig3_df, var = "ndi", var_label = "Neighborhood Deprivation Index", var_breaks = 2) # NDI
figs2_svi <- gen_fig3_scatterplot(datain = fig3_df, var = "svi", var_label = "Social Vulnerability Index", var_breaks = 0.2) # SVI
### EXPORT: Plots
fig3_ndi
ggsave('figs/fig3.png', width = 15, height = 12)
figs2_svi
ggsave('figs/figs2.png', width = 15, height = 12)


```

## FIGURE 4 - PART B (CLUSTER MAP + CLUSTER PROFILE)
```{r}
## CHECK WHICH CLUSTERS TRACTS WERE IN
tract_final %>% filter(Tract_FIPS %in% c('36047051900','36081067100')) %>% select(Tract_FIPS, borough, nevi, nevi_cluster, score_demo, score_economic, score_residential, score_healthstatus) # Queens tract: 4, Kings tract: 6
#View where on map: https://popfactfinder.planning.nyc.gov/#11.95/40.72281/-73.95688
# library(leaflet)
# tract_final_spatial_tempproj <- st_transform(tract_final_spatial, crs = '+proj=longlat +datum=WGS84')
# tract_final_spatial_tempproj_subset <- tract_final_spatial_tempproj %>% filter(Tract_FIPS %in% c('36047051900','36081067100'))
# leaflet() %>%
#   addTiles() %>%
#   addPolygons(data = tract_final_spatial_tempproj, color = "blue", label = tract_final$Tract_FIPS,
#               popup = c(paste0('GEOID: ', tract_final$Tract_FIPS))) %>%
#   addPolygons(data = tract_final_spatial_tempproj %>% filter(Tract_FIPS %in% c('36047051900','36081067100')), color = "red", label = tract_final_spatial_tempproj$Tract_FIPS,
#               popup = c(paste0('GEOID: ', tract_final$Tract_FIPS)))

## CREATE: PART 1 - CLUSTER PROFILES
# Colors
# Explore color options
  # colorBlindness::displayAllColors(paletteMartin, color="white")
  # colors_colorblindsafe <- paletteMartin[c(5, 6, 9, 11, 14, 15)] %>% unname()
colors_colorblindsafe <- c('#ffff6d','#ffb6db','#24ff24','#6db6ff','#920000','#490092') 

# Alt: colors_colorblindsafe <- safeColors[2:7] %>% unname()
# Ref: https://www.r-graph-gallery.com/circular-barplot.html
nevi_cat_levels <- c('Demographics', 'Economic', 'Residential', 'Health Status')
# Data Prep
fig4_part1_df <- tract_final %>% 
  dplyr::group_by(nevi_cluster) %>% 
  dplyr::summarize(n_obs = n(),
                   mean_nevi = mean(nevi),
                   q1_nevi = quantile(nevi, 0.25),
                   q3_nevi = quantile(nevi, 0.75),
                   iqr_nevi = paste0(round(q1_nevi, 2), "-", round(q3_nevi,2)),
                   mean_demo_score = mean(score_demo),
                   mean_economic_score = mean(score_economic),
                   mean_residential_score = mean(score_residential),
                   mean_healthstatus_score = mean(score_healthstatus)) %>% 
  tidyr::gather(nevi_cat_prelim, mean, mean_demo_score:mean_healthstatus_score) %>% 
  dplyr::mutate(nevi_cat = case_when(nevi_cat_prelim == 'mean_demo_score'         ~ nevi_cat_levels[1],
                                    nevi_cat_prelim == 'mean_economic_score'     ~ nevi_cat_levels[2],
                                    nevi_cat_prelim == 'mean_residential_score'  ~ nevi_cat_levels[3],
                                    nevi_cat_prelim == 'mean_healthstatus_score' ~ nevi_cat_levels[4]) %>% factor(nevi_cat_levels),
                nevi_cluster_label = paste0("Cluster ", nevi_cluster, " (n = ", n_obs, ",\nMean NEVI = ", round(mean_nevi,2), ")"))
# Plotting
fig4_part1_prelim <- ggplot(fig4_part1_df, aes(x = nevi_cat, y = mean, fill = nevi_cluster_label, pattern = nevi_cat)) +
  ggpattern::geom_bar_pattern(stat = "identity", color = "black") + # settings if legend patterns do not show: pattern_density = 0.3, pattern_key_scale_factor = 0.6
  geom_text(aes(label = round(mean, 2), y = mean + 0.25), size = 4) + # y = mean + 0.5
  coord_polar(start = (3*pi)/2, direction = -1, clip = "off") +
  ylim(-0.2, 1.1) + #1.1
  scale_fill_manual(values = colors_colorblindsafe, labels = c("1","2","3","4","5","6")) +
  ggpattern::scale_pattern_manual(values = c("circle","stripe","crosshatch","none")) +
  facet_wrap(. ~ nevi_cluster_label, nrow = 2) +
  labs(fill = "Neighborhood\nEnvironmental\nVulnerability Index\nCluster", pattern = "Domain Score") +
  theme_minimal() +
  guides(fill = guide_legend(override.aes = list(pattern = "none"), nrow = 2, title.position = "top", title.hjust = 0.5, order = 1), # fill = FALSE 
         pattern = guide_legend(override.aes = list(fill = "white"), nrow = 6, title.position = "top", title.hjust = 0.5, order = 2)) +
  theme(legend.position = "right",
        legend.title = element_text(face = "bold"),
        legend.box = 'vertical',
        axis.title = element_blank(),
        axis.text = element_blank(),
        strip.text.x = element_text(size = 12, face = "bold"),
        panel.grid = element_blank(),
        plot.margin = unit(c(t = 0, r = 0, b = 0, l = 0), "cm"),
        panel.spacing = unit(0, "lines")) # strip.background.y = element_rect("transparent")
fig4_part1_plot_build <- ggplot_build(fig4_part1_prelim)
for(i in 1:6){
  fig4_part1_plot_build[["layout"]][["panel_params"]][[i]][["r.range"]][2] <- 0.45
}
fig4_part1 <- ggplot_gtable(fig4_part1_plot_build)


## CREATE: PART 2 - CLUSTER MAP
gen_fig4_part2 <- function(dfin, dfexcl, dfborough, fill_var, fill_color, fill_label, fill_increment){
  plot_choro <- ggplot() + 
    geom_sf(data = dfin, aes(fill = !!sym(fill_var)), color = NA) + 
    geom_sf(data = dfexcl, color = "gray90", fill = "gray90") + 
    geom_sf(data = dfborough, color = 'gray20', fill = NA, size = 0.6) +
    theme_minimal() + 
    scale_fill_manual(values = fill_color, guide = guide_legend(direction = "horizontal", title.position = "top")) +
    labs(fill = fill_label) + 
    # guides(fill = guide_legend(nrow = 1, title.position = "top", title.hjust = 0.5)) +
    theme(legend.position = "none", # legend.title = element_text(face = "bold", size = 12), legend.text = element_text(size = 9), legend.spacing.x = unit(0.5, 'cm'), 
          axis.line = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank(),
          panel.grid = element_blank(),
          plot.margin = unit(c(t = 0, r = 0, b = 0, l = 0), "cm")) 
  return(plot_choro)
}
fig4_part2 <- gen_fig4_part2(dfin = tract_final_spatial, dfexcl = tract_exclude_spatial, dfborough = borough_nyc_spatial, fill_var = "nevi_cluster", fill_color = colors_colorblindsafe, fill_label = "Neighborhood Environmental Vulnerability Index Cluster") 

## COMBINED PLOT
fig4b <- ggarrange(fig4_part1, ggarrange(fig4_part2, ggplot() + theme_void(), ncol = 2), nrow = 2) + theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0, "cm"))
fig4b
# EXPORT
ggsave('figs/fig4_part.png', width = 7, height = 9) # pixels: 1200 x 900 width = 13, height = 12

# View average NEVI domains scores by cluster for text
fig4b_part1_df %>% distinct(nevi_cluster, mean_nevi, iqr_nevi)

```


### FIGURE 5: HEATMAP
```{r}
### CREATE: Heat maps to visualize clustering by subdomain
## PREPARE DATA
subdomain_label_vector <- c(
  'Age','Female-Led Households','Immigration','Disability','Single Parent Households','Transportation','Single Households',
  'Income and Poverty','Occupation','Income Inequality','Employment Status','Education','Vehicle Availability',
  'Population Density','Group Quarters','Household Density','Age of Housing Structure','Units in Housing Structure','Residential Mobility','Vacancy',
  'Health Behaviors','Health Conditions','Prevention Practices','Health Insurance Status'
)
domain_label_vector <- c('Demographic\n Domain', 'Economic Indicators\n Domain', 'Residential\n Domain', 'Health Status\n Domain')
fig5_df <- stab5 %>% 
  dplyr::as_tibble(rownames = 'subdomain') %>% 
  dplyr::filter(!row_number() %in% c(1,2,3)) %>% 
  dplyr::transmute(subdomain = subdomain,
                   cluster1 = str_sub(V1, 1L, 4L) %>% as.numeric,
                   cluster2 = str_sub(V2, 1L, 4L) %>% as.numeric,
                   cluster3 = str_sub(V3, 1L, 4L) %>% as.numeric,
                   cluster4 = str_sub(V4, 1L, 4L) %>% as.numeric,
                   cluster5 = str_sub(V5, 1L, 4L) %>% as.numeric,
                   cluster6 = str_sub(V6, 1L, 4L) %>% as.numeric,
                   subdomain_label = case_when(
                     subdomain == 'score_demo_age_median_iqr' ~ subdomain_label_vector[1],
                     subdomain == 'score_demo_femaleled_median_iqr' ~ subdomain_label_vector[2],
                     subdomain == 'score_demo_immigration_median_iqr' ~ subdomain_label_vector[3],
                     subdomain == 'score_demo_disability_median_iqr' ~ subdomain_label_vector[4],
                     subdomain == 'score_demo_singleparent_median_iqr' ~ subdomain_label_vector[5],
                     subdomain == 'score_demo_mobility_median_iqr' ~ subdomain_label_vector[6],
                     subdomain == 'score_demo_livealone_median_iqr' ~ subdomain_label_vector[7],
                     subdomain == 'score_economic_incomepoverty_median_iqr' ~ subdomain_label_vector[8],
                     subdomain == 'score_economic_servicemanual_median_iqr' ~ subdomain_label_vector[9],
                     subdomain == 'score_economic_gini_median_iqr' ~ subdomain_label_vector[10],
                     subdomain == 'score_economic_employment_median_iqr' ~ subdomain_label_vector[11],
                     subdomain == 'score_economic_education_median_iqr' ~ subdomain_label_vector[12],
                     subdomain == 'score_economic_vehicleavail_median_iqr' ~ subdomain_label_vector[13],
                     subdomain == 'score_residential_popdensity_median_iqr' ~ subdomain_label_vector[14],
                     subdomain == 'score_residential_groupquarters_median_iqr' ~ subdomain_label_vector[15],
                     subdomain == 'score_residential_occperroom_median_iqr' ~ subdomain_label_vector[16],
                     subdomain == 'score_residential_structage_median_iqr' ~ subdomain_label_vector[17],
                     subdomain == 'score_residential_structattach_median_iqr' ~ subdomain_label_vector[18],
                     subdomain == 'score_residential_move1yr_median_iqr' ~ subdomain_label_vector[19],
                     subdomain == 'score_residential_vacancy_median_iqr' ~ subdomain_label_vector[20],
                     subdomain == 'score_healthstatus_lifestyle_median_iqr' ~ subdomain_label_vector[21],
                     subdomain == 'score_healthstatus_condition_median_iqr' ~ subdomain_label_vector[22],
                     subdomain == 'score_healthstatus_preventive_median_iqr' ~ subdomain_label_vector[23],
                     subdomain == 'score_healthstatus_lackinsurance_median_iqr' ~ subdomain_label_vector[24]
                   ) %>% factor(levels = subdomain_label_vector, ordered = TRUE), 
                   domain = case_when(
                     stringr::str_detect(subdomain, 'demo') ~ domain_label_vector[1],
                     stringr::str_detect(subdomain, 'economic') ~ domain_label_vector[2],
                     stringr::str_detect(subdomain, 'residential') ~ domain_label_vector[3],
                     stringr::str_detect(subdomain, 'healthstatus') ~ domain_label_vector[4]
                   ) %>% factor(levels = domain_label_vector, ordered = TRUE)) %>%  
  dplyr::select(-subdomain) %>% 
  tidyr::pivot_longer(cols = starts_with('cluster'), names_to = 'cluster', names_prefix = 'cluster', values_to = 'median') %>% 
  dplyr::mutate(cluster_label = case_when(
    cluster == '1' ~ '1\n(n = 435)',
    cluster == '2' ~ '2\n(n = 331)',
    cluster == '3' ~ '3\n(n = 903)',
    cluster == '4' ~ '4\n(n = 5)',
    cluster == '5' ~ '5\n(n = 241)',
    cluster == '6' ~ '6\n(n = 171)'
  )) %>% 
  dplyr::group_by(subdomain_label) %>% 
  dplyr::mutate(median_scaled = scale(median))

##### CREATE: heatmap 
ggplot(data = fig5_df, aes(x = cluster_label, y = subdomain_label, fill = median_scaled)) +
  geom_tile() + 
  facet_grid(domain ~ ., scales = 'free_y', space = 'free_y') +
  scale_y_discrete(limits = rev) +
  scale_fill_gradient(low = 'white', high = 'red') +
  labs(x = 'NEVI Cluster', y = 'Subdomain', fill = "Median Score (Scaled)") +
  theme_minimal() +
  theme(plot.background = element_rect(fill = 'white', color = 'white'),
        axis.title.x = element_text(size = 20, margin = margin(t = 10)), axis.text.x = element_text(size = 15),
        axis.title.y = element_text(size = 20, margin = margin(t = 5)), axis.text.y = element_text(size = 15), 
        strip.text = element_text(size = 16),
        legend.title = element_text(size = 16), legend.text = element_text(size = 14),
        legend.position = 'top', legend.key.width = unit(1.5, 'cm'), legend.spacing.x = unit(1, 'cm'))
## EXPORT
ggsave('figs/fig5.png', width = 10, height = 13)
```


## SUPPLEMENTAL FIGURE 1. Box plots
```{r}
# note changed colors for overall and residential
figs1_df <- tract_final %>% 
  dplyr::select(Tract_FIPS, borough, nevi, score_demo, score_economic, score_residential, score_healthstatus) %>% 
  tidyr::gather(key = index_type, value = index, nevi, score_demo, score_economic, score_residential, score_healthstatus) %>% 
  dplyr::mutate(index_type = case_when(index_type == "nevi" ~ "Overall:\nNEVI",
                                       index_type == "score_demo" ~ "Domain:\nDemographic\nScore",
                                       index_type == "score_economic" ~ "Domain:\nEconomic\nScore",
                                       index_type == "score_residential" ~ "Domain:\nResidential\nScore",
                                       index_type == "score_healthstatus" ~ "Domain:\nHealth Status\nScore") %>% 
                                       factor(levels = c("Overall:\nNEVI","Domain:\nDemographic\nScore","Domain:\nEconomic\nScore","Domain:\nResidential\nScore","Domain:\nHealth Status\nScore",
                                                         'Neighborhood Deprivation Index','Social Vulnerability Index')),
                part = 'NEVI') %>% 
  rbind(ndi_clean %>% dplyr::mutate(index_type = 'Neighborhood Deprivation Index', part = 'NDI') %>% rename(index = ndi),
        svi_clean %>% dplyr::mutate(index_type = 'Social Vulnerability Index', part = 'SVI') %>% rename(index = svi)) 
### CREATE: Plots
  # NEVI
figs1a <- ggplot(figs1_df %>% filter(part == 'NEVI'), aes(y = borough, x = index, fill = index_type)) +
  geom_boxplot() +
  facet_grid(index_type ~ ., scale = "free") +
  # labs(x = "Index", y = "Borough") +
  labs(title = "Neighborhood Environmental Vulnerability Index") +
  scale_fill_manual(values = c("#D55E00","#0072B2","#009E73","#E69F00","#CC79A7")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_blank(),  axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), strip.text.y = element_text(size = 20), plot.title = element_text(face = "bold", hjust = 0.5, size = 20))
  # NDI
figs1b <- ggplot(figs1_df %>% filter(part == 'NDI'), aes(y = borough, x = index, fill = index_type)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#D55E00")) +
  labs(title = "Neighborhood Deprivation Index") +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_blank(),  axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), plot.title = element_text(face = "bold", hjust = 0.5, size = 20))
  # SVI
figs1c <- ggplot(figs1_df %>% filter(part == 'SVI'), aes(y = borough, x = index, fill = index_type)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#D55E00")) +
  labs(title = "Social Vulnerability Index") +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_blank(),  axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), plot.title = element_text(face = "bold", hjust = 0.5, size = 20))
### EXPORT: Plots
# fig3_labs <- c("A","B","C") # c("Neighborhood Environmental Vulnerability Index","Neighborhood Deprivation Index","Social Vulnerability Index")
figs1 <- ggarrange(figs1a, figs1b, figs1c, nrow = 3, font.label = list(size = 25), align = "v", heights = c(4,1,1)) #labels = fig3_labs,
annotate_figure(figs1,left = text_grob("Borough", rot = 90, hjust = 1, size = 20), bottom = text_grob("Index", size = 20))
ggsave('figs/figs1.png', width = 15, height = 16)
```

### SUPPLEMENTAL FIGURE 3 Sensitivity analysis: Race and ethnicity
```{r}
### Clean data 
# CREATE: Function
clean_toxpi_sa <- function(datain, databorough){
  dataout <- datain %>% 
    dplyr::rename(Tract_FIPS = Source, nevi = `ToxPi Score`) %>% 
    dplyr::mutate(Tract_FIPS = as.character(Tract_FIPS) %>% trimws()) %>% 
    dplyr::select(-c(`HClust Group`, `KMeans Group`, Name)) %>% 
    # Demographics
    dplyr::rename_at(vars(starts_with('DemographicsAge')), funs(paste0('score_demo_age'))) %>% 
    dplyr::rename_at(vars(starts_with('DemographicsFemaleLed')), funs(paste0('score_demo_femaleled'))) %>% 
    dplyr::rename_at(vars(starts_with('DemographicsRaceEth!')), funs(paste0('score_demo_raceeth'))) %>% 
    dplyr::rename_at(vars(starts_with('DemographicsImmigration')), funs(paste0('score_demo_immigration'))) %>% 
    dplyr::rename_at(vars(starts_with('DemographicsDisability')), funs(paste0('score_demo_disability'))) %>% 
    dplyr::rename_at(vars(starts_with('DemographicsSingleParent')), funs(paste0('score_demo_singleparent'))) %>% 
    dplyr::rename_at(vars(starts_with('DemographicsMobility')), funs(paste0('score_demo_mobility'))) %>% 
    dplyr::rename_at(vars(starts_with('DemographicsLiveAlone')), funs(paste0('score_demo_livealone'))) %>% 
    # Economic
    dplyr::rename_at(vars(starts_with('Economic IndicatorsIncomePoverty')), funs(paste0('score_economic_incomepoverty'))) %>% 
    dplyr::rename_at(vars(starts_with('Economic IndicatorsServiceManualJobs')), funs(paste0('score_economic_servicemanual'))) %>% 
    dplyr::rename_at(vars(starts_with('Economic IndicatorsGini')), funs(paste0('score_economic_gini'))) %>% 
    dplyr::rename_at(vars(starts_with('Economic IndicatorsEmployment')), funs(paste0('score_economic_employment'))) %>% 
    dplyr::rename_at(vars(starts_with('Economic IndicatorsEducation')), funs(paste0('score_economic_education'))) %>% 
    dplyr::rename_at(vars(starts_with('Economic IndicatorsVehicleAvail')), funs(paste0('score_economic_vehicleavail'))) %>% 
    # Residential
    dplyr::rename_at(vars(starts_with('Residential DensityPopDensity')), funs(paste0('score_residential_popdensity'))) %>% 
    dplyr::rename_at(vars(starts_with('Residential DensityGroupQuarters')), funs(paste0('score_residential_groupquarters'))) %>% 
    dplyr::rename_at(vars(starts_with('Residential DensityOccPerRoom')), funs(paste0('score_residential_occperroom'))) %>% 
    dplyr::rename_at(vars(starts_with('Residential DensityStrAge')), funs(paste0('score_residential_structage'))) %>% 
    dplyr::rename_at(vars(starts_with('Residential DensityStrAttached')), funs(paste0('score_residential_structattach'))) %>% 
    dplyr::rename_at(vars(starts_with('Residential DensityMove1Yr')), funs(paste0('score_residential_move1yr'))) %>% 
    dplyr::rename_at(vars(starts_with('Residential DensityVacancy')), funs(paste0('score_residential_vacancy'))) %>% 
    # Health status
    dplyr::rename_at(vars(starts_with('Health StatusLifestyle')), funs(paste0('score_healthstatus_lifestyle'))) %>% 
    dplyr::rename_at(vars(starts_with('Health StatusCondition')), funs(paste0('score_healthstatus_condition'))) %>% 
    dplyr::rename_at(vars(starts_with('Health StatusPreventive')), funs(paste0('score_healthstatus_preventive'))) %>% 
    dplyr::rename_at(vars(starts_with('Health StatusLackInsurance')), funs(paste0('score_healthstatus_lackinsurance'))) %>% 
    dplyr::mutate(score_economic = (score_economic_incomepoverty + score_economic_servicemanual + score_economic_gini + score_economic_employment + score_economic_education + score_economic_vehicleavail)/6,
                  score_residential = (score_residential_popdensity + score_residential_groupquarters + score_residential_occperroom + score_residential_structage + score_residential_structattach + score_residential_move1yr + score_residential_vacancy)/7,
                  score_healthstatus = (score_healthstatus_lifestyle + score_healthstatus_condition + score_healthstatus_preventive + score_healthstatus_lackinsurance)/4) %>% 
    dplyr::left_join(databorough, by = "Tract_FIPS")
  return(dataout)
}
# DATA CLEANING
  # Race/eth: no overlapping
tract_toxpi_raceeth_no_olap_df <- clean_toxpi_sa(tract_toxpi_pre_raceeth_no_olap, tract_nyc_borough) %>% 
  dplyr::mutate(score_demo = (score_demo_age + score_demo_femaleled + score_demo_raceeth + score_demo_immigration + score_demo_disability + score_demo_singleparent + score_demo_mobility + score_demo_livealone)/8) 

##### PLOTS
### BOXPLOT
# PREP data
figs3_df <- rbind(tract_final %>% dplyr::select(nevi, score_demo, borough) %>% dplyr::mutate(index_type = "Without Race and Ethnicity \nCategories (Main)"),
               tract_toxpi_raceeth_no_olap_df %>% dplyr::select(nevi, score_demo, borough) %>% dplyr::mutate(index_type = "With Race and Ethnicity\nCategories")) %>% 
  dplyr::mutate(all = "All Boroughs", index_type = factor(index_type, levels = c("Without Race and Ethnicity \nCategories (Main)",
                                                           "With Race and Ethnicity\nCategories")))

# CREATE: Plots
  # NEVI
figs3a <- ggplot(figs3_df, aes(x = nevi, fill = index_type)) +
  geom_boxplot(aes(y = borough)) +
  geom_boxplot(aes(y = all)) +
  scale_y_discrete(limits = rev) +
  geom_hline(yintercept = 5.5, color = "gray20", linetype = "dashed") +
  labs(title = "Overall NEVI", fill = "NEVI:\nRace and Ethnicity Categorization") +
  scale_fill_manual(values = c("#D55E00", "gray40", "gray60", "gray80")) + # limits = rev
  guides(fill = guide_legend(reverse = TRUE)) +
  theme_bw() +
  theme(axis.title.x = element_blank(),  axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), panel.grid.major.y = element_blank(), strip.text.y = element_text(size = 20), plot.title = element_text(face = "bold", hjust = 0.5, size = 20), legend.title = element_text(size = 16), legend.text = element_text(size = 14))
  # Domain: Demographic 
figs3b <- ggplot(figs3_df, aes(x = score_demo, fill = index_type)) +
  geom_boxplot(aes(y = borough)) +
  geom_boxplot(aes(y = all)) +
  scale_y_discrete(limits = rev) +
  geom_hline(yintercept = 5.5, color = "gray20", linetype = "dashed") +
  labs(title = "Domain: Demographic Score", fill = "Demographic Domain:\nRace and Ethnicity Categorization") +
  scale_fill_manual(values = c("#0072B2", "gray40", "gray60", "gray80")) +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme_bw() +
  theme(axis.title.x = element_blank(),  axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), panel.grid.major.y = element_blank(), plot.title = element_text(face = "bold", hjust = 0.5, size = 20), legend.title = element_text(size = 16), legend.text = element_text(size = 14))
### MERGE AND EXPORT: Plots
figs3 <- ggarrange(figs3a, figs3b, nrow = 2, font.label = list(size = 25), align = "v", heights = c(1,1,1)) +
  theme(plot.margin=unit(c(0,0,0.2,0.2),"in"),
        panel.background = element_rect(fill = "white", color = "white"))

figs3_export <- annotate_figure(figs3,left = text_grob("Borough", rot = 90, hjust = 1, size = 20), bottom = text_grob("Index", size = 20))
ggsave('figs/figs3.png', plot = figs3_export, width = 10, height = 10)

```

## SUPPLEMENTAL FIGURE 4 - RACE AND ETHNICITY CLUSTER MAP + PROFILE (SIMILAR TO FIG4)
```{r}
## CREATE: Clusters: NEVI with non-overlapping race and ethnicity features
set.seed(123)
  # Get slice weights
w_raceeth <- sapply(sapply(names(tract_toxpi_pre_raceeth_no_olap)[-c(1:5)],function(x) strsplit(x, "!")),"[",2)
w_raceeth <- sapply(strsplit(w_raceeth,split="/"),function(x) {y <- ifelse(length(x)==2,x[2],1);as.numeric(x[1])/as.numeric(y)})
  # Generate cluster
hc_raceeth <- hclust(dist(tract_toxpi_pre_raceeth_no_olap[,-c(1:5)]*rep(w_raceeth,each=nrow(tract_toxpi_pre_raceeth_no_olap))), method="complete")
nevi_cluster_result_raceeth <- cutree(hc_raceeth, k = 6) # Cut into 6 clusters using HClusts 
  # Create data frame for cluster data
tract_nevi_cluster_raceeth <- tract_toxpi_pre_raceeth_no_olap %>% 
  dplyr::mutate(Tract_FIPS = Source %>% as.character() %>% trimws(),
                nevi_cluster_orig = nevi_cluster_result_raceeth %>% as.factor(),
                nevi_cluster = case_when(
                  nevi_cluster_orig == "1" ~ "6",
                  nevi_cluster_orig == "2" ~ "5",
                  nevi_cluster_orig == "3" ~ "4",
                  nevi_cluster_orig == "4" ~ "3",
                  nevi_cluster_orig == "5" ~ "1",
                  nevi_cluster_orig == "6" ~ "2"
                ) %>% as.factor()) %>% 
  dplyr::select(Tract_FIPS, nevi_cluster)
  # Merge with cleaned df
figs4_df <- tract_toxpi_raceeth_no_olap_df %>% 
  dplyr::left_join(tract_nevi_cluster_raceeth, by = 'Tract_FIPS')
figs4_spatial <- tract_nyc_spatial %>% 
  dplyr::inner_join(figs4_df %>% dplyr::mutate(Tract_FIPS = as.character(Tract_FIPS) %>% trimws()), by = "Tract_FIPS")

## CHECK CLUSTERS
# scaled_tract_toxpi_pre_raceeth_no_olap = scale(tract_toxpi_pre_raceeth_no_olap[-c(1:5)])
# mydist <-function(x)dist(x, method="euclidian")
# mycluster <- function(x,k) list(cluster=cutree(hclust(mydist(x), method = "complete"),k=k))
# gap_stat_10_raceeth_no_olap <- clusGap(scaled_tract_toxpi_pre_raceeth_no_olap, FUN = mycluster, K.max = 10, B = 50)
# library(factoextra)
# fviz_gap_stat(gap_stat_10_raceeth_no_olap) # choose 6 clusters based on this plot and what we observed when characterizing the clusters.



## CREATE: PART 1 - CLUSTER PROFILES
# Colors
colors_colorblindsafe <- c('#ffff6d','#ffb6db','#6db6ff','#24ff24','#920000','#490092') 

# Alt: colors_colorblindsafe <- safeColors[2:7] %>% unname()
# Ref: https://www.r-graph-gallery.com/circular-barplot.html
nevi_cat_levels <- c('Demographics', 'Economic', 'Residential', 'Health Status')
# Data Prep
figs4_part1_df <- figs4_df %>% 
  dplyr::group_by(nevi_cluster) %>% 
  dplyr::summarize(n_obs = n(),
                   mean_nevi = mean(nevi),
                   q1_nevi = quantile(nevi, 0.25),
                   q3_nevi = quantile(nevi, 0.75),
                   iqr_nevi = paste0(round(q1_nevi, 2), "-", round(q3_nevi,2)),
                   mean_demo_score = mean(score_demo),
                   mean_economic_score = mean(score_economic),
                   mean_residential_score = mean(score_residential),
                   mean_healthstatus_score = mean(score_healthstatus)) %>% 
  tidyr::gather(nevi_cat_prelim, mean, mean_demo_score:mean_healthstatus_score) %>% 
  dplyr::mutate(nevi_cat = case_when(nevi_cat_prelim == 'mean_demo_score'         ~ nevi_cat_levels[1],
                                    nevi_cat_prelim == 'mean_economic_score'     ~ nevi_cat_levels[2],
                                    nevi_cat_prelim == 'mean_residential_score'  ~ nevi_cat_levels[3],
                                    nevi_cat_prelim == 'mean_healthstatus_score' ~ nevi_cat_levels[4]) %>% factor(nevi_cat_levels),
                nevi_cluster_label = paste0("Cluster ", nevi_cluster, " (n = ", n_obs, ",\nMean NEVI = ", round(mean_nevi,2), ")"))
# Plotting
figs4_part1_prelim <- ggplot(figs4_part1_df, aes(x = nevi_cat, y = mean, fill = nevi_cluster_label, pattern = nevi_cat)) +
  ggpattern::geom_bar_pattern(stat = "identity", color = "black") +
  geom_text(aes(label = round(mean, 2), y = mean + 0.25), size = 4) + # y = mean + 0.5
  coord_polar(start = (3*pi)/2, direction = -1, clip = "off") +
  ylim(-0.2, 1.1) + #1.1
  scale_fill_manual(values = colors_colorblindsafe, labels = c("1","2","3","4","5","6")) +
  ggpattern::scale_pattern_manual(values = c("circle","stripe","crosshatch","none")) +
  facet_wrap(. ~ nevi_cluster_label, nrow = 2) +
  labs(fill = "Neighborhood\nEnvironmental\nVulnerability Index\nCluster", pattern = "Domain Score") +
  theme_minimal() +
  guides(fill = guide_legend(override.aes = list(pattern = "none"), nrow = 2, title.position = "top", title.hjust = 0.5, order = 1), # fill = FALSE 
         pattern = guide_legend(override.aes = list(fill = "white"), nrow = 6, title.position = "top", title.hjust = 0.5, order = 2)) +
  theme(legend.position = "right",
        legend.title = element_text(face = "bold"),
        legend.box = 'vertical',
        axis.title = element_blank(),
        axis.text = element_blank(),
        strip.text.x = element_text(size = 12, face = "bold"),
        panel.grid = element_blank(),
        plot.margin = unit(c(t = 0, r = 0, b = 0, l = 0), "cm"),
        panel.spacing = unit(0, "lines")) # strip.background.y = element_rect("transparent")
figs4_part1_plot_build <- ggplot_build(figs4_part1_prelim)
for(i in 1:6){
  figs4_part1_plot_build[["layout"]][["panel_params"]][[i]][["r.range"]][2] <- 0.45
}
figs4_part1 <- ggplot_gtable(figs4_part1_plot_build)


## CREATE: PART 2 - CLUSTER MAP
gen_figs4_part2 <- function(dfin, dfexcl, dfborough, fill_var, fill_color, fill_label, fill_increment){
  plot_choro <- ggplot() + 
    geom_sf(data = dfin, aes(fill = !!sym(fill_var)), color = NA) + 
    geom_sf(data = dfexcl, color = "gray90", fill = "gray90") + 
    geom_sf(data = dfborough, color = 'gray20', fill = NA, size = 0.6) +
    theme_minimal() + 
    scale_fill_manual(values = fill_color, guide = guide_legend(direction = "horizontal", title.position = "top")) +
    labs(fill = fill_label) + 
    # guides(fill = guide_legend(nrow = 1, title.position = "top", title.hjust = 0.5)) +
    theme(legend.position = "none", # legend.title = element_text(face = "bold", size = 12), legend.text = element_text(size = 9), legend.spacing.x = unit(0.5, 'cm'), 
          axis.line = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank(),
          panel.grid = element_blank(),
          plot.margin = unit(c(t = 0, r = 0, b = 0, l = 0), "cm")) 
  return(plot_choro)
}
figs4_part2 <- gen_figs4_part2(dfin = figs4_spatial, dfexcl = tract_exclude_spatial, dfborough = borough_nyc_spatial, fill_var = "nevi_cluster", fill_color = colors_colorblindsafe, fill_label = "Neighborhood Environmental Vulnerability Index Cluster") 

## COMBINED PLOT
figs4b <- ggarrange(figs4_part1, figs4_part2, nrow = 2) + theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0, "cm"), panel.background = element_rect(fill = 'white', color = 'white'))
figs4b
# EXPORT
ggsave('figs/figs4.png', width = 7, height = 9) # pixels: 1200 x 900 width = 13, height = 12

```


### FIGURE 5: HEATMAP
```{r}
### CREATE: Heat maps to visualize clustering by subdomain
## PREPARE DATA
subdomain_label_vector <- c(
  'Age','Female-Led Households','Immigration','Disability','Single Parent Households','Transportation','Single Households',
  'Income and Poverty','Occupation','Income Inequality','Employment Status','Education','Vehicle Availability',
  'Population Density','Group Quarters','Household Density','Age of Housing Structure','Units in Housing Structure','Residential Mobility','Vacancy',
  'Health Behaviors','Health Conditions','Prevention Practices','Health Insurance Status'
)
domain_label_vector <- c('Demographic\n Domain', 'Economic Indicators\n Domain', 'Residential\n Domain', 'Health Status\n Domain')
fig5_df <- stab5 %>% 
  dplyr::as_tibble(rownames = 'subdomain') %>% 
  dplyr::filter(!row_number() %in% c(1,2,3)) %>% 
  dplyr::transmute(subdomain = subdomain,
                   cluster1 = str_sub(V1, 1L, 4L) %>% as.numeric,
                   cluster2 = str_sub(V2, 1L, 4L) %>% as.numeric,
                   cluster3 = str_sub(V3, 1L, 4L) %>% as.numeric,
                   cluster4 = str_sub(V4, 1L, 4L) %>% as.numeric,
                   cluster5 = str_sub(V5, 1L, 4L) %>% as.numeric,
                   cluster6 = str_sub(V6, 1L, 4L) %>% as.numeric,
                   subdomain_label = case_when(
                     subdomain == 'score_demo_age_median_iqr' ~ subdomain_label_vector[1],
                     subdomain == 'score_demo_femaleled_median_iqr' ~ subdomain_label_vector[2],
                     subdomain == 'score_demo_immigration_median_iqr' ~ subdomain_label_vector[3],
                     subdomain == 'score_demo_disability_median_iqr' ~ subdomain_label_vector[4],
                     subdomain == 'score_demo_singleparent_median_iqr' ~ subdomain_label_vector[5],
                     subdomain == 'score_demo_mobility_median_iqr' ~ subdomain_label_vector[6],
                     subdomain == 'score_demo_livealone_median_iqr' ~ subdomain_label_vector[7],
                     subdomain == 'score_economic_incomepoverty_median_iqr' ~ subdomain_label_vector[8],
                     subdomain == 'score_economic_servicemanual_median_iqr' ~ subdomain_label_vector[9],
                     subdomain == 'score_economic_gini_median_iqr' ~ subdomain_label_vector[10],
                     subdomain == 'score_economic_employment_median_iqr' ~ subdomain_label_vector[11],
                     subdomain == 'score_economic_education_median_iqr' ~ subdomain_label_vector[12],
                     subdomain == 'score_economic_vehicleavail_median_iqr' ~ subdomain_label_vector[13],
                     subdomain == 'score_residential_popdensity_median_iqr' ~ subdomain_label_vector[14],
                     subdomain == 'score_residential_groupquarters_median_iqr' ~ subdomain_label_vector[15],
                     subdomain == 'score_residential_occperroom_median_iqr' ~ subdomain_label_vector[16],
                     subdomain == 'score_residential_structage_median_iqr' ~ subdomain_label_vector[17],
                     subdomain == 'score_residential_structattach_median_iqr' ~ subdomain_label_vector[18],
                     subdomain == 'score_residential_move1yr_median_iqr' ~ subdomain_label_vector[19],
                     subdomain == 'score_residential_vacancy_median_iqr' ~ subdomain_label_vector[20],
                     subdomain == 'score_healthstatus_lifestyle_median_iqr' ~ subdomain_label_vector[21],
                     subdomain == 'score_healthstatus_condition_median_iqr' ~ subdomain_label_vector[22],
                     subdomain == 'score_healthstatus_preventive_median_iqr' ~ subdomain_label_vector[23],
                     subdomain == 'score_healthstatus_lackinsurance_median_iqr' ~ subdomain_label_vector[24]
                   ) %>% factor(levels = subdomain_label_vector, ordered = TRUE), 
                   domain = case_when(
                     stringr::str_detect(subdomain, 'demo') ~ domain_label_vector[1],
                     stringr::str_detect(subdomain, 'economic') ~ domain_label_vector[2],
                     stringr::str_detect(subdomain, 'residential') ~ domain_label_vector[3],
                     stringr::str_detect(subdomain, 'healthstatus') ~ domain_label_vector[4]
                   ) %>% factor(levels = domain_label_vector, ordered = TRUE)) %>%  
  dplyr::select(-subdomain) %>% 
  tidyr::pivot_longer(cols = starts_with('cluster'), names_to = 'cluster', names_prefix = 'cluster', values_to = 'median') %>% 
  dplyr::mutate(cluster_label = case_when(
    cluster == '1' ~ '1\n(n = 435)',
    cluster == '2' ~ '2\n(n = 331)',
    cluster == '3' ~ '3\n(n = 903)',
    cluster == '4' ~ '4\n(n = 5)',
    cluster == '5' ~ '5\n(n = 241)',
    cluster == '6' ~ '6\n(n = 171)'
  )) %>% 
  dplyr::group_by(subdomain_label) %>% 
  dplyr::mutate(median_scaled = scale(median))

##### CREATE: heatmap 
ggplot(data = fig5_df, aes(x = cluster_label, y = subdomain_label, fill = median_scaled)) +
  geom_tile() + 
  facet_grid(domain ~ ., scales = 'free_y', space = 'free_y') +
  scale_y_discrete(limits = rev) +
  scale_fill_gradient(low = 'white', high = 'red') +
  labs(x = 'NEVI Cluster', y = 'Subdomain', fill = "Median Score (Scaled)") +
  theme_minimal() +
  theme(plot.background = element_rect(fill = 'white', color = 'white'),
        axis.title.x = element_text(size = 20, margin = margin(t = 10)), axis.text.x = element_text(size = 15),
        axis.title.y = element_text(size = 20, margin = margin(t = 5)), axis.text.y = element_text(size = 15), 
        strip.text = element_text(size = 16),
        legend.title = element_text(size = 16), legend.text = element_text(size = 14),
        legend.position = 'top', legend.key.width = unit(1.5, 'cm'), legend.spacing.x = unit(1, 'cm'))
## EXPORT
ggsave('figs/fig5.png', width = 10, height = 13)



### SUPPLEMENTAL TABLE 1 & INFO FOR TEXT
```{r}
# Prep data
stab1_df <- tract_final %>% 
  dplyr::left_join(ndi_clean, by = "Tract_FIPS") %>% 
  dplyr::left_join(svi_clean, by = "Tract_FIPS")


# Function to generate summary stats
gen_stab1 <- function(dfin, var, var_label){
  var <- sym(var)
  overall <- dfin %>% 
    dplyr::summarize(mean = mean(!!var), sd = sd(!!var), median = median(!!var), min = min(!!var), max = max(!!var), q1 = quantile(!!var, probs = 0.25), q3 = quantile(!!var, probs = 0.75), iqr = paste0(q1 %>% round(2) %>% format(nsmall = 2),"-",q3 %>% round(2) %>% format(nsmall = 2)), median_iqr = paste0(median %>% round(2) %>% format(nsmall = 2)," (IQR = ",iqr,")"), median_iqr2 = paste0(median %>% round(2) %>% format(nsmall = 2),", IQR = ",iqr), median_iqr3 = paste0("median = ", median_iqr2)) %>% 
    dplyr::mutate(borough = "Overall")
  byborough <- dfin %>% 
    dplyr::group_by(borough) %>% 
    dplyr::summarize(mean = mean(!!var), sd = sd(!!var), median = median(!!var), min = min(!!var), max = max(!!var), q1 = quantile(!!var, probs = 0.25), q3 = quantile(!!var, probs = 0.75), iqr = paste0(q1 %>% round(2) %>% format(nsmall = 2),"-",q3 %>% round(2) %>% format(nsmall = 2)), median_iqr = paste0(median %>% round(2) %>% format(nsmall = 2)," (IQR = ",iqr,")"), median_iqr2 = paste0(median %>% round(2) %>% format(nsmall = 2),", IQR = ",iqr), median_iqr3 = paste0("median = ", median_iqr2)) %>% 
    dplyr::mutate(borough = paste0("   ", borough))
  combined <- rbind(overall, byborough) %>% 
    dplyr::mutate(score = var_label) %>% 
    dplyr::relocate(score, borough, mean, sd, median, iqr, min, max, median_iqr, median_iqr2, median_iqr3, .before = mean)
  return(combined)
}

# Generate summary stats
stab1_full <- rbind(gen_stab1(stab1_df, "nevi", "Neighborhood Environmental Vulnerability Index (NEVI)")
                   ,gen_stab1(stab1_df, "ndi", "Neighborhood Deprivation Index (NDI)")
                   ,gen_stab1(stab1_df, "svi", "Social Vulnerability Index (SVI)")
                   ,gen_stab1(stab1_df, "score_demo", "NEVI Domain: Demographics")
                   ,gen_stab1(stab1_df, "score_economic", "NEVI Domain: Economic Indicators")
                   ,gen_stab1(stab1_df, "score_residential", "NEVI Domain: Residential Characteristics")
                   ,gen_stab1(stab1_df, "score_healthstatus", "NEVI Domain: Health Status"))

# text_summary %>% View()
export(stab1_full, 'tabs/stab1_full.xlsx')

# OVERALL CORRELATIONS
cor.test(stab1_df$nevi, stab1_df$ndi, method = "spearman")
cor.test(stab1_df$nevi, stab1_df$svi, method = "spearman")


```


### SUPPLEMENTAL TABLE 5: Cluster summaries
```{r}
# Function to generate summary stats
gen_stab4 <- function(dfin, var){
  var <- sym(var)
  var_median <- sym(paste0('median_iqr_', var))
  nevi_cluster_summary <- dfin %>% 
    dplyr::group_by(nevi_cluster) %>% 
    dplyr::summarize(iqr_prelim = quantile(!!var, c(0.25, 0.75)) %>% round(2) %>% format(2)) %>% 
    dplyr::group_by(nevi_cluster) %>% 
    dplyr::summarize(iqr = paste(iqr_prelim, collapse = '-'))
  tab_summary <- dfin %>% 
    dplyr::group_by(nevi_cluster) %>%
    dplyr::summarize(median_nevi = median(!!var) %>% round(2) %>% format(nsmall = 2)) %>% 
    dplyr::left_join(nevi_cluster_summary, by = 'nevi_cluster') %>% 
    dplyr::summarize(nevi_cluster = nevi_cluster,
                     median_iqr = paste0(median_nevi,' (',iqr,')'))
  return(tab_summary)
}
# Generate summary stats
stab4 <- tract_final %>% 
  janitor::tabyl(nevi_cluster) %>% 
  janitor::adorn_pct_formatting(digits = 2) %>% 
  dplyr::transmute(nevi_cluster = nevi_cluster,
                   n_percent = paste0('(n = ', n,', ', percent,')')) %>% 
  dplyr::left_join(gen_stab4(tract_final, 'nevi') %>% dplyr::rename(nevi_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_demo_age') %>% dplyr::rename(score_demo_age_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_demo_femaleled') %>% dplyr::rename(score_demo_femaleled_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_demo_immigration') %>% dplyr::rename(score_demo_immigration_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_demo_disability') %>% dplyr::rename(score_demo_disability_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_demo_singleparent') %>% dplyr::rename(score_demo_singleparent_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_demo_mobility') %>% dplyr::rename(score_demo_mobility_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_demo_livealone') %>% dplyr::rename(score_demo_livealone_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_economic_incomepoverty') %>% dplyr::rename(score_economic_incomepoverty_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_economic_servicemanual') %>% dplyr::rename(score_economic_servicemanual_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_economic_gini') %>% dplyr::rename(score_economic_gini_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_economic_employment') %>% dplyr::rename(score_economic_employment_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_economic_education') %>% dplyr::rename(score_economic_education_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_economic_vehicleavail') %>% dplyr::rename(score_economic_vehicleavail_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_residential_popdensity') %>% dplyr::rename(score_residential_popdensity_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_residential_groupquarters') %>% dplyr::rename(score_residential_groupquarters_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_residential_occperroom') %>% dplyr::rename(score_residential_occperroom_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_residential_structage') %>% dplyr::rename(score_residential_structage_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_residential_structattach') %>% dplyr::rename(score_residential_structattach_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_residential_move1yr') %>% dplyr::rename(score_residential_move1yr_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_residential_vacancy') %>% dplyr::rename(score_residential_vacancy_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_healthstatus_lifestyle') %>% dplyr::rename(score_healthstatus_lifestyle_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_healthstatus_condition') %>% dplyr::rename(score_healthstatus_condition_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_healthstatus_preventive') %>% dplyr::rename(score_healthstatus_preventive_median_iqr = median_iqr)) %>%
  dplyr::left_join(gen_stab4(tract_final, 'score_healthstatus_lackinsurance') %>% dplyr::rename(score_healthstatus_lackinsurance_median_iqr = median_iqr)) %>% 
  t()
## EXPORT
export(stab4, 'tabs/stab4.xlsx', row.names = TRUE)

```






