---
title: "Neighborhood Environmental Vulnerability Index, 2019: Calculate Zip Code-Level NEVI"
author: 
- "Stephen P. Uong; Contributors: Jiayi Zhou, Jeanette A. Stingone"
date: "3/29/2022"
output: rmarkdown::github_document
---

### 1. Set Working Directory

Set the working directory to one folder up from the RMarkdown file for later data export.

```{r, setup}
knitr::opts_knit$set(root.dir = '..') 
```

### 2. Load Required Libraries

Load the following required libraries.

```{r, libraries}
library(tidyverse)
library(rio)
```

### 3. Import the Data

Import the following data files:

-   [Modified Zip Code Tabulation Areas (MODZCTA) used by the New York City Department of Health & Mental Hygiene (NYC DOHMH)](https://data.cityofnewyork.us/Health/Modified-Zip-Code-Tabulation-Areas-MODZCTA-/pri4-ifjk)

    -   We previously downloaded this data in the link above and saved the file in `data/raw/Modified Zip Code Tabulation Areas`

-   [U.S. Department of Housing and Urban Development - Zip Code and Tract Crosswalk](https://www.huduser.gov/portal/datasets/usps_crosswalk.html)

    -   We previously downloaded this data in the link above and saved the file in `data/raw/US Department of Housing and Urban Development - Crosswalk`

```{r, import_data}
# Zip code list used by NYC DOHMH ZCTA list
list_nyc_zip <- readr::read_csv('data/raw/NYC Modified Zip Code Tabulation Areas/Modified_Zip_Code_Tabulation_Areas__MODZCTA_.csv') %>% 
  dplyr::select(ZCTA) %>% 
  tidyr::separate_rows(ZCTA, sep = ',') %>% 
  dplyr::transmute(zip = trimws(ZCTA))
# U.S. Department of Housing and Urban Development - Zip Code and Tract Crosswalk
xwalk_zip_tract <- readxl::read_excel("data/raw/US Department of Housing and Urban Development - Crosswalk/ZIP_TRACT_122019.xlsx", col_types = c(rep('guess',2),rep('numeric',4)))
# Tract-level NEVI scores generate from the generate_nevi.Rmd code
nevi_tract <- readRDS(file = "data/processed/nevi_tract_final.rds")
```

### 4. Inclusion and Exclusion

We included zip codes (ZCTAs) used by the NYC DOHMH and excluded zip codes that had a residential ratio of 0.

```{r}
# Include: NYC DOHMH zip codes only
xwalk_zip_tract_nyc_prelim <- xwalk_zip_tract %>%
  dplyr::select(ZIP, TRACT, RES_RATIO) %>% 
  dplyr::rename(zip = ZIP) %>% 
  dplyr::mutate(zip = zip %>% trimws(),
                TRACT = TRACT %>% trimws()) %>% 
  dplyr::inner_join(list_nyc_zip, by = 'zip')
# Exclusion list: Generate list of zip codes with total residential ratio = 0 (30 NYC zip codes)
list_exclude_zip_resratio0 <- xwalk_zip_tract_nyc_prelim %>% 
  dplyr::group_by(zip) %>% 
  dplyr::summarize(RES_RATIO_total = sum(RES_RATIO)) %>% 
  dplyr::filter(RES_RATIO_total == 0)
# Exclude: zip codes with total residential ratio = 0
xwalk_zip_tract_nyc <- xwalk_zip_tract_nyc_prelim %>% 
  dplyr::anti_join(list_exclude_zip_resratio0, by = 'zip')
```

### 5. Calculate Zip Code-Level NEVI

Recalculate the NEVI from tracts to zip codes with the following steps:

-   Multiplied the tract-level scores by the proportion of residential addresses in a given zip code within the Census tracts (residential ratio).

-   Summed up the scores weighted by the residential ratios by zip code.

-   Divided these summed scores by the summed residential ratios (because some zip codes did not fully overlap with NYC and therefore had a total residential ratio of \< 1).

```{r}
# Keep only columns needed to recalculate NEVI from tract to zip code
nevi_tract_subset <- nevi_tract %>% 
  dplyr::rename(TRACT = Tract_FIPS) %>% 
  dplyr::select(TRACT, nevi, score_demo, score_economic, score_residential, score_healthstatus)
# Calculate by zip code
nevi_zip_calc <- xwalk_zip_tract_nyc %>%
  dplyr::inner_join(nevi_tract_subset, by = 'TRACT') %>% 
  # Multiply tract-level scores by residential ratio
  dplyr::transmute(zip = zip,
                   tract = TRACT,
                   res_ratio_part = RES_RATIO,
                   nevi_part = nevi*RES_RATIO,
                   score_demo_part = score_demo*RES_RATIO,
                   score_economic_part = score_economic*RES_RATIO,
                   score_residential_part = score_residential*RES_RATIO,
                   score_healthstatus_part = score_healthstatus*RES_RATIO) %>% 
  dplyr::group_by(zip) %>% 
  # Sum scores by zip code (multiple tracts per zip code)
  dplyr::summarize(res_ratio_sum = sum(res_ratio_part),
                   nevi = sum(nevi_part),
                   score_demo = sum(score_demo_part),
                   score_economic = sum(score_economic_part),
                   score_residential = sum(score_residential_part),
                   score_healthstatus = sum(score_healthstatus_part)) %>% 
  # Divide summed scores by summed residential ratio (b/c not all tracts in crosswalk file was in the NEVI tract dataset)
  dplyr::mutate(nevi = nevi/res_ratio_sum,
                score_demo = score_demo/res_ratio_sum,
                score_economic = score_economic/res_ratio_sum,
                score_residential = score_residential/res_ratio_sum,
                score_healthstatus = score_healthstatus/res_ratio_sum) %>% 
  # EXCLUDE: 1 Zip (11430) with total RES_RATIO = 0 (b/c other census tract in zip code was not in census tract dataset, excluded b/c pop <20 or missing feature)
  dplyr::filter(res_ratio_sum != 0) %>% 
  dplyr::select(-res_ratio_sum)
```

### 7. Export Zip Code-Level NEVI

```{r}
export(nevi_zip_calc, 'data/processed/nevi_zip_final.csv')
saveRDS(nevi_zip_calc, file = "data/processed/nevi_zip_final.rds")
```
